# **Introduction to Containers**

Before we start to address your project with Collab O. Rator, we will provide an introduction to {{< glossary "Containerization" display="containerization" >}} with {{< glossary "Docker" >}}. In this section, you will learn more about what containers are, why they are useful, and when they may be unnecessary. In additional you will learn about different software tools or technologies that support containerization, when certain technologies are more appropriate to use, as well as how to find, share, and work with containers. Towards the end of this section we will return to the {{< glossary "Microbiome" display="microbiome" >}}-{{< glossary "Metabolome" display="metabolome" >}} data collection project.

## What is a container?

A [container is an active computing environment](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-a-container/) that is isolated from the rest of the user's system. Containers are the operational ending point of the containerization process (involving Dockerfiles and images), define and instantiate or produce containers. 

You can think of containers like the countryside world from the Mary Poppins scene where Mary Poppins, Bert, and the kids jump from a London park into the chalk drawing of the countryside. The chalk drawing is the "Image" that is "run" to launch the running container. 

```{r, echo = FALSE, fig.alt = "Images are a frozen snapshot of what the computing environment will be in a container or what will be in the running container, while a container is the isolated, specified computing environment that you can work in. This is like the Mary Poppins scene where Mary Poppins, Bert, and the kids are looking at chalk drawings (images) and jump into one which has its own isolated environment (container)."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f020ee26d5_2_0#slide=id.g3f020ee26d5_2_0")
```

Containers and the containerization process are supported by different tools such as Docker, Podman, Apptainer, and Singularity. Docker is one of the industry-standard and most popular of the tools and primarily will be the one discussed and utilized within this case study. 

We will discuss the containerization process that leads to containers, starting at the beginning:

### Dockerfiles

Dockerfiles are the starting point in the containerization process. Dockerfiles are text-based scripts that contain the commands that will be used to provision the eventual container computing environment. These commands will be evaluated sequentially one after the other when the Dockerfile is "built" to produce an image. In this way, Dockerfiles are a blueprint, specifying what the eventual container will have available in its computing environment for the user. Dockerfiles are small and relatively easy to store and read. Because they are text-based, they can be version-controlled. The sorts of [commands that will be in Dockerfiles](https://docs.docker.com/get-started/docker-concepts/building-images/writing-a-dockerfile/) include speciyfing a base image, which dependencies or software packages should be installed, a file structure to set up, defining persistent environment variables. 

```{r, echo = FALSE, fig.alt = "Dockerfiles are the starting point in the containerization process "}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f3ca37d253_0_0#slide=id.g3f3ca37d253_0_0")
```

For other tools (e.g., Singularity or Apptainer) the equivalent of these files are instead called definition files.

### Building

"Build" is the verb that is used to move from a Dockerfile to an image within the containerization process. Dockerfiles are built to produce images. Images are built in a way that the resulting containers will be able to run on specific platforms or architectures (e.g., hardware + operating system: linux or windows). [Multi-platform builds](https://docs.docker.com/build/building/multi-platform/) allow for even greater flexibility in who can use a container no matter what hardware + operating system they have. For example, newer Macs have a different processor chip (Apple silicon) than older macs (Intel) and may have more difficulty building, pulling, or running older images. [See this description about known issues](https://docs.docker.com/desktop/troubleshoot-and-support/troubleshoot/known-issues/) for ways around this. 

### Images

[Images are layered, frozen snapshots](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-an-image/) that contain all of the necessary files, configurations, package libraries, etc that are necessary to produce the active running container. An image is not an active environment, but rather an intermediate step in the containerization process. Each layer is related to a command from the Dockerfile: adding or modifying to what the container will contain. Once images are created they are immutable (e.g., cannot be changed or modified). If you need something to be changed, you would have to create a new image. You may want to use the same image name but a different "tag" if it is a small change or updated version of an old/no longer necessary image. Common tags include "main", "latest", "dev", "v#", etc. Because images are immutable and layered, you can use images as bases for other images, starting with an existing image and adding a package or set of packages on top of it. Or removing a package because it's no longer needed and you can have a smaller image. Images are much larger than Dockerfiles (often Gigabytes (GB) in size) and not as readable. Images are the main way that containers are shared. 

```{r, echo = FALSE, fig.alt = "Images are layered and built by processing the sequential commands in a Dockerfile, stacking each layer on the previous layer according to the commands in the Dockerfile."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f01a9d8250_0_17#slide=id.g3f01a9d8250_0_17")
```

### Registries

Images can be stored within your local file system but you might want to free up space (due to their large size) or share them with team members, collaborators, another machine you use, or because you are publishing a study. This is where [registeries](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-a-registry/) are useful. A registry is a centralized location for storing and sharing container images. Docker Hub is a common, but not the only, registry used to share Docker images.

```{r, echo = FALSE, fig.alt = "Images are stored in registries and can be pushed or pulled from a registry. The images that are pushed to a registry are built from a Dockerfile. The images that are pulled from the registry will match the images that were pushed. If you pull an image and later use the original Dockerfile to build an image, these could be the same image or they might be different if undefined package or dependency versions have changed."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3fb50cd5a7c_0_323#slide=id.g3fb50cd5a7c_0_323")
```

Images can be "pushed" or "pulled" from a registry. If an image has multiple tags or are built for multiple architectures, the related images (with different tags or built for different architectures) will be stored in a [repository within the registry](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-a-registry/#registry-vs-repository).  If an image contains proprietary information, it likely shouldn't be shared in a public registry like Docker Hub.

An image that is pulled from a registry will match the image that was pushed to the registry. If after some time (after pushing the image to the registry) an image is built from the Dockerfile that was used to build the original image (that was pushed to the registry), these images may no longer match. Unless versions are explicitly controlled within the Dockerfile, versions of dependencies or packages may have changes or may no longer be available. Also, you may be using different options or parameters to build the Docker image than were originally used. Small changes like this can impact how reproducible a container actually is in practice and is a reason why Images are typically stored in registries to share. 

### Running

"Run" is the verb that is used to move from an image to a container within the containerization process. Images are run to create or launch a container -- the active, usable computing environment. Once a container is running, files from your system cannot be added to the container because the container is isolated. You can likely use [`wget`](https://www.geeksforgeeks.org/linux-unix/wget-command-in-linux-unix/) to fetch files from the internet, but those files will not be available to others using the same image to run another version of the container unless they fetch them as well. When you "run" an image to create a container you can mount a volume of your local files so that files from your system are available to the container. These files may be a analysis scripts, data files, package that you are writing/developingAny changes made to those files will still be available to you once the container is no longer running.

```{r, echo = FALSE, fig.alt = "Volumes have to be mounted when running an image to have access to outside files such as scripts or data."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f613e4255f_0_0#slide=id.g3f613e4255f_0_0")
```

If your instiution has strict security configurations set on your machine (possibly due to sensitive data), you may not be able to mount volumes when running a Docker container. 

<!--
Discuss daemon, root level permissions, and principle of least privilege here? 
-->

::: {.column-margin}
Find alternatives to suggest for these cases. I think using Podman is one we can mention, but are there others for working with Docker?
:::


::: {.emphasis_block}
In summary, Dockerfiles are text-based scripts that contain commands defining how an image will be built. Dockerfiles serve as a blueprint or the starting point of the containerization process. Dockerfiles are built to produce images. Images are a frozen, layered snapshot of what the container will be. Images are much larger than Dockerfiles and less readable. They are the most common way to share a container and are often stored in registries (like Docker Hub). Images are run to create containers and file systems or volumes may be mounted during the run process so that isolated containers have access to scripts or data and any updates to those are available outside of the container as well.
:::

::: {.emphasis_block}

<details><summary><b>Containerization Analogies</b></summary>

::: {.panel-tabset}

### Kitchen and cooking analogy

The containerization process can be thought of like working in a kitchen for dinner service or a holiday meal. 

```{r fig.align = 'center', out.width = '100%', echo = FALSE, fig.alt = 'Diagram showing how working in a kitchen can be an analogy for the containerization process. Dockerfiles specify the recipe repertoire and kitchen organization standards while building is like the preparation process resulting in standardized resources, ready for active cooking, but not actively being used - an image. Running the image results in a container or an active kitchen where cooking is happening. It is like turning on the burner. In the running process you may mount a volume of scripts and data or in the kitchen this would be specific recipes. Finally, container clean up is removing the heat and resetting the environment by storing ingredients and cleaning up the space.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f88b06ed70_0_0#slide=id.g3f88b06ed70_0_0")
```

In the kitchen you have a recipe or a set of recipes each listing the necessary ingredients to stock as well as how much of each ingredient is needed and if there's any special preparation or considerations for an ingredient (e.g., chopping or peeling). Recipes also provide detailed instructions for the environment (desired pans, cooking temperature, time). Additionally, a kitchen has specific locations for specific tools, pans, and utensils -- perhaps there's a map or binder specifying where things are located and how they are to be cleaned and stored. All of these together specify how the kitchen should be provisioned.

The Dockerfile is like a recipe and kitchen binder specifying what dependencies, packages, and environmental variables (ingredients) are needed, in what order, and what the environment needs to look like. While a Dockerfile contains a layered, step-by-step definition of the computing environment, it doesn't contain a step-by-step definition of how a specific analysis will utilize the computing environment. 

Working in a kitchen, often the area and ingredients may be prepped before any combining, cooking, or baking actually happens. Pans, utensils, and tools are all cleaned and stored in specific, known locations. Ingredients that need peeled, chopped, or some other special preparation are prepared in those ways. This preparation allows for a more efficient or easier combination of ingredients within the cooking process. Everything is prepared (according to kitchen station expectations and the instructions for ingredients in the recipe), just sitting at the ready, but not actively being used. This is termed "mise en place" (pronunciation: /ˌmiz ɑn ˈplɑs/ or meez ahn PLAHSS).

The "build" process (Dockerfile --> image) is the preparation of the environment and ingredients: the cleaning, stocking, chopping, dicing, peeling, measuring, etc. which leads to the docker image where everything is in a static, read-only, prepared state. The docker image is like "mise en place".

The step-by-step instructions for specific dishes, and their utilization of specific ingredients may be needed as well (analysis scripts). Perhaps unusual ingredients (data) for special dishes are needed. These are not part of the building process, but instead are brought into the kitchen (container) once everything else is ready. The recipes, kitchen station and prep guidelines (Dockerfile), and prepared environment, and ingredients (image) provide a static and standardized setup -- one that can be used by any set of cooks or chefs. The mounted volume within a running container brings in the flexibility to use the static and standardized setup for your specific needs in your specific way.

Once the kitchen is ready, they fire up a stove top, heat an oven, or combine ingredients that will work together in a bowl. With the addition of heat or mixing of ingredients, the cooking process itself begins. Everything that was defined in the recipe or kitchen station instructions is at the kitchen's disposal. It's too late to go get different ingredients or supplies.

The "run" process (image --> container) is like supplying the heat -- creating the environment where the cooking and combining of ingredients can happen -- the container. In the running container or environment, each of the ingredients (packages) are at your disposal, ready to use. Your analysis scripts and specific data will be incorporated here by mounting a volume during the "run" process. 

Within the running container, you use the packages together with any data or analysis scripts to perform an analysis or provision a tool. Everything defined in the Dockerfile is at your disposal to use in the running container. And the running container is isolated from the rest of your computer -- so only what is defined in the Dockerfile is available unless you've mounted a volume to bring in specific data and scripts.

The eventual meals produced by the kitchen are like the finished product: a tool, analysis, result, etc. that can be shared. And finally, as with any kitchen endeavour -- there has to be a clean up. Perhaps a container is paused or stopped (e.g., the heat is turned off but the pan is still available on the stove). Then eventually the container is removed -- (e.g., dishes are cleared, cleaned, and put away). Images may also be removed to save space -- like clearing out or storing unused ingredients. 

In a kitchen, you may make multiple courses for a meal: an appetizer, the main course, and dessert. Each of these have their own process: own station in a professional kitchen, own set of recipes, own prep needs, and own tool needs. However, the end product is the whole meal. In a similar way, an analysis may have multiple parts each supported by its own container, all working together to produce the overall end product.

### Hotel analogy

The containerization process can be thought of like hotels. 

```{r fig.align = 'center', out.width = '100%', echo = FALSE, fig.alt = 'Diagram showing how the containerization process can be thought of as hotels.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f8b4188434_0_21#slide=id.g3f8b4188434_0_21")
```

Add an analogy about hotel rooms

### Lego analogy

The containerization process can be thought of like building with Legos.

```{r fig.align = 'center', out.width = '100%', echo = FALSE, fig.alt = 'Diagram showing how the containerization process can be thought of as building legos.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f8b4188434_0_21#slide=id.g3f8b4188434_0_21")
```

Add an analogy about building legos

:::

</details>

:::

## Why containers?

As previously described, containers are usefufl for standardizing the computing environment within a team or between different machines (e.g., local laptop and institutional HPC or cloud computing environment).

Containers are not always necessary. The complexity of setting up, sharing, and documeting containers may not be worth what you gain in regards to reproducibility. In such cases reporting software used and their versions is often enough. Such cases include: 

* If you're working on an anlysis without specialized software. Specialized software is software that may be hard to install because of dependencies or software that you wouldn't expect many people in your field to work with.
* If you're **not** working within a large collaborative team.
* If an analysis can stay in one place.
* If the computing platform already supports analyses through the use of containers (e.g., The NHGRI AnVIL Platform), then you may not need to set up your own container if theirs already provides the software you need.

### When is a container especially useful?

Containers are especially useful in the following cases:

* If you need to move an analysis to another machine or environment like an HPC or the cloud
* If a collaborative team is working on a large analysis within a workflow (e.g., WDL, Snakemake) or in different environments
* If a collaborative team is having difficulty replicating results within the team
* If dependencies are causing nightmares

 --> Discuss Singularity and Apptainer for HPCs in this section <br>
  --> Include why Docker is not appropriate for HPCs <br>
  --> Recommendations for the cloud <br>

### What are general use cases for containers?

There are several different use cases for containers: 

* Supporting an analysis
* Deploying an app

### What are the first steps for working with a container?

* Developing: starting from the beginning, setting up a base image, specifying dependencies, actual packages of interest
* Finding: search and find a container that meets your needs (you may have to modify it a bit/add a couple other packages)
* Specified: A collaborator sends you a container and you need to use it (or add a package or otherwise modify it in some way to suit your needs)

The last of these scenarios is the context for this case study.

```{r, fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = ''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f020ee26d5_4_12#slide=id.g3f020ee26d5_4_12")
```

## How do you work with a container?

### Options for interacting with the container

How do you work with a container and run an analysis with a container?

* Terminal / Command Line
  * If you only have a Dockerfile, you need to use this to {{< glossary "Build" display="build" >}} the image
* Docker Desktop
* An interactive Integrated Development Environment (IDE)
* In combination with GitHub Actions

Sometimes the process of running a container and interacting with it involves a combination of these options (because the container installs an IDE) that once you've run the container, you then develop an analysis within an IDE.

#### Docker Desktop vs Docker through a Command Line Interface (CLI)

The reason we're primarily walking learners through using Docker Desktop is because

1. It separates out images and containers which will support learners trying to learn the difference
2. Learners can create files, as well as import and export files within the files tab as opposed to mounting a volume
3. Learners can also mount a volume which is an option
4. Learners can easily see how many images and containers they have and while they are learning this can be extremely helpful so they don't tax their systems too much
5. Docker Desktop has its own documentation that we can point to