<!--
More details on the background on the case study topic as well its relation to biomedical science.
Videos or images may be helpful so this includes an example of embedding a video.
-->

# **Context**
***

## Metabolomics and Data Resources

Discuss information such as ...

* why the microbiome and metabolome datasets are related/considered together
* what the study with the table we're looking at is analyzing or presenting to the scientific community
* what a cohort is
* what paired data means
* what longitudinal means
* what HMDB-annotated compounds are
* what KEGG-annotated compounds are
* possible uses of this data resource


## The need for containerization

Data analysis, especially analysis of large data, relies on computers and the software tools installed on those copmuters. Depending on the computational needs of an analysis or a researcher's typical work, that computer or work station may be a personal computer / laptop (Operating system: Microsoft Windows, Apple MacOS, Google ChromeOS, or Linux), a desktop, a Graphics Processing Unit (GPU) server, shared computing resources such as institutional high performance computing (HPC) clusters or supercomputers, or the cloud (Amazon Web Services, Google Cloud Platform, Microsoft Azure). [@Wright_Computing].

Different researchers, especially those located at different institutions, those at different career stages, and those who have different research focuses have different computing environments (potentially differences in operating systems, software, software versions, etc.). Beyond differences in operating systems, different institutions and the IT departments that manage the institution's computers may have specific setups and guidelines shaping the available software and computing environment for computers at the institution. Individuals with different needs will install and update software differentially. Even team members within the same lab will have different computing environments. Overtime, a single researcher's computer will evolve as well. Software and operating systems will be updated (sometimes automatically rather than manually) and new software, tools, or applications will be installed. Basically, computers are like snowflakes -- they are unique when compared to other computers. They make have similarities, but overall computers are their own unique entity.

```{r fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = 'Within a team or collaborative environment, each computer is unique – like snowflakes.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f6b98e19f5_0_1#slide=id.g3f6b98e19f5_0_1")
```

The differences between computers can be problematic for performing and sharing reproducible research. 


```{r fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = 'Containers provide a consistent computing environment for everyone on the team, isolated from the unique setups.'}
ottrpal::include_slide('https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f6b98e19f5_0_49#slide=id.g3f6b98e19f5_0_49')
```

***
