# **Introduction to Containers**

Before we start to address your project with Collab O. Rator, we will provide an introduction to containerization with Docker. In this section, you will learn more about what containers are, why they are useful, and when they may be unnecessary. In additional you will learn about different software tools or technologies that support containerization, when certain technologies are more appropriate to use, as well as how to find, share, and work with containers. Towards the end of this section we will return to the microbiome-metabolome data collection project.

## What is a container?

* Describe images, containers, and other vocabulary around the subject such as Dockerfile, building an image, etc.
* make a diagram showing how Dockerfile relates to building an image ... (possibly borrow or adapt the crane one that Candace made for ITCR containers course)

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f020ee26d5_2_0#slide=id.g3f020ee26d5_2_0")
```

The containerization process can be thought of like working in a kitchen for dinner service or holiday meal. 

In the kitchen you have a recipe or a set of recipes each listing the necessary ingredients to stock as well as how much of each ingredient is needed and if there's any special preparation or considerations for an ingredient (e.g., chopping or peeling). Recipes also provide detailed instructions for the environment (desired pans, cooking temperature, time). Additionally, a kitchen has specific locations for specific tools, pans, and utensils -- perhaps there's a map or binder specifying where things are located and how they are to be cleaned and stored. All of these together specify how the kitchen should be provisioned.

The Dockerfile is like a recipe and kitchen binder specifying what dependencies, packages, and environmental variables (ingredients) are needed, in what order, and what the environment needs to look like. While a Dockerfile contains a layered, step-by-step definition of the computing environment, it doesn't contain a step-by-step definition of how a specific analysis will utilize the computing environment. 

Working in a kitchen, often the area and ingredients may be prepped before any combining, cooking, or baking actually happens. Pans, utensils, and tools are all cleaned and stored in specific, known locations. Ingredients that need peeled, chopped, or some other special preparation are prepared in those ways. This preparation allows for a more efficient or easier combination of ingredients within the cooking process. Everything is prepared (according to kitchen station expectations and the instructions for ingredients in the recipe), just sitting at the ready, but not actively being used. This is termed "mise en place".

The "build" process (Dockerfile --> image) is the preparation of the environment and ingredients: the cleaning, stocking, chopping, dicing, peeling, measuring, etc. which leads to the docker image where everything is in a static, read-only, prepared state. The docker image is like "mise en place".

The step-by-step instructions for specific dishes, and their utiliziation of specific ingredients may be needed as well (analysis scripts). Perhaps unusual ingredients (data) for special dishes are needed. These are not part of the building process, but instead are brought into the container once everything else is ready. The recipes, kitchen station and prep guidelines (Dockerfile), and prepared environment, and ingredients (image) provide a static and standardized setup -- one that can be used by any set of cooks or chefs. The mounted volume within a running container brings in the flexibility to use the static and standardized setup for your specific needs in your specific way.

Once the kitchen is ready, they fire up a stovetop, heat an oven, or combine ingredients that will work together in a bowl. With the addition of heat or mixing of ingredients, the cooking process itself begins. Everything that was defined in the recipe or kitchen station instructions is at the kitchen's disposal. It's too late to go get different ingredients or supplies.

The "run" process (image --> container) is like supplying the heat -- creating the environment where the cooking and combining of ingredients can happen -- the container. In the running container or environment, each of the ingredients (packages) are at your disposal, ready to use. Your analysis scripts and specific data will be incorporated here by mounting a volume during the "run" process. 

Within the running container, you use the packages together with any data or analysis scripts to perform an analysis or provision a tool. Everything defined in the Dockerfile is at your disposal to use in the running container. And the running container is isolated from the rest of your computer -- so only what is defined in the Dockerfile is available unless you've mounted a volume to bring in specific data and scripts.

The eventual meals produced by the kitchen are like the finished product: a tool, analysis, result, etc. that can be shared. And finally, as with any kitchen endeavour -- there has to be a clean up. Perhaps a container is paused or stopped (e.g., the heat is turned off but the pan is still available on the stove). Then eventually the container is removed -- (e.g., dishes are cleared, cleaned, and put away). Images may also be removed to save space -- like clearing out or storing unused ingredients. 

In a kitchen, you may make multiple courses for a meal: an appetizer, the main course, and dessert. Each of these have their own process: own station in a professional kitchen, own set of recipes, own prep needs, and own tool needs. However, the end product is the whole meal. In a similar way, an analysis may have multiple parts each supported by its own container, all working together to produce the overall end product.

## Why containers?

* Describe why
* Describe that a container isn't always necessary, but more on this later...


### What are general use cases for containers?

* Developing: starting from the beginning, setting up a base image, specifying dependencies, actual packages of interest
* Finding: search and find a container that meets your needs (you may have to modify it a bit/add a couple other packages)
* Specified: A collaborator sends you a container and you need to use it (or add a package or otherwise modify it in some way to suit your needs)

The last of these scenarios is the context for this case study.

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f020ee26d5_4_12#slide=id.g3f020ee26d5_4_12")
```

### When is a container especially useful?

* Especially if you need to move an analysis to another machine or environment like an HPC or the cloud

  --> Discuss Singularity and Apptainer for HPCs in this section <br>
  --> Include why Docker is not appropriate for HPCs <br>
  --> Recommendations for the cloud <br>

* If a collaborative team is working on a large analysis within a workflow or in different environments
* If a collaborative team is having difficulty replicating results within the team
* If dependencies are causing nightmares

### When might a container be unnecessary?

When is a container overkill? Do you really need a container or will session info work?

## Where can you find and share containers?

... more on this within the data import section but highlight Dockerhub, Github, and someone like a collaborator sending you a container here... re point to the email

## How do you work with a container?

### Options for interacting with the container

How do you work with a container and run an analysis with a container?

* Terminal / Command Line
  * If you only have a Dockerfile, you need to use this to build the image
* Docker Desktop
* An interactive Integrated Development Environment (IDE)
* In combination with GitHub Actions

Sometimes the process of running a container and interacting with it involves a combination of these options (because the container installs an IDE) that once you've run the container, you then develop an analysis within an IDE.

We'll primarily walk learners through using Docker Desktop though we'll likely use a tab structure to give learners the optional methods to use Terminal / Command Line. We will also provide a side quest about using a container that installs an IDE and suggest an advanced homework for the final option of combining containers with GitHub Actions.

The reason we're primarily walking learners through using Docker Desktop is because

1. It separates out images and containers which will support learners trying to learn the difference
2. Learners can create files, as well as import and export files within the files tab as opposed to mounting a volume
3. Learners can also mount a volume which is an option
4. Learners can easily see how many images and containers they have and while they are learning this can be extremely helpful so they don't tax their systems too much
5. Docker Desktop has its own documentation that we can point to

### Options for interacting with data within a running container

As for how we're working with data. Our options are to

1. mount a volume (directory) from our system that has data (and possibly scripts) in it.
2. `cp` files to and from the container
3. Create an analysis script within the running container and edit it within the container. Use a URL to get the data PDF from within the container and create/keep raw data and intermediates within the container that won't be shared. Finally, creating plots within the container that we'll have to save to our system.

While we will show examples of all of these, our primary interest is to work within the third option as much as possible. When we ran an event at a medical center, the devices were unable to mount data so we'd like to provide options separate from that as much as possible.