<!--
More details on the background on the case study topic as well its relation to biomedical science.
Videos or images may be helpful so this includes an example of embedding a video.
-->

# **Context**
***

## Metabolomics and Data Resources

Discuss information such as ...

* why the microbiome and metabolome datasets are related/considered together
* what the study with the table we're looking at is analyzing or presenting to the scientific community
* what a cohort is
* what paired data means
* what longitudinal means
* what HMDB-annotated compounds are
* what KEGG-annotated compounds are
* possible uses of this data resource

## The revision task -- Extracting data from a PDF and data visualization

While this example of data only being available from a PDF (especially within an active project) is a bit unconventional, we think that it will be a tangible and somewhat accessible example for learners. Previous public health case studies highlight various data mining and import tasks such as utilizing an API, scraping data from an HTML webpage, and importing data from Excel files, Google Sheets, and even PDFs. This specific package hasn't been used in previous case studies and it offers some benefits for this specific topic

* This package has a function for reading tables from a PDF file. Several of the previously used tools do not, and the one that does is deprecated.
* This package has some dependencies (namely Java) that we are excited to highlight within the context of containers and showing a tiny bit of complexity within the Dockerfile
* Its use helps to justify to the learners why we might need a container (dependencies)
* We don't intend to spend much of the main narrative on the tool or how to use it. Some of the explanation can go into collapsible sections to reduce cognitive load so that learners can focus on containers. Basically we'll tell them that the data retrieval and wrangling of the Table from a PDF can be achieved with this code and if you want to learn more about how it works (or want an advanced exercise of generalizing the process), click on an expandable section for details.
* We can note that reducing manual steps and automating things (like data retrieval rather than manually creating a spreadsheet from visually inspecting the table) are important steps towards more reproducible analyses because they are less error prone and more transparent.

## The need for containerization

Data analysis, especially analysis of large data, relies on computers and the software tools installed on those copmuters. Depending on the computational needs of an analysis or a researcher's typical work, that computer or work station may be a personal computer / laptop (Operating system: Microsoft Windows, Apple MacOS, Google ChromeOS, or Linux), a desktop, a Graphics Processing Unit (GPU) server, shared computing resources such as institutional high performance computing (HPC) clusters or supercomputers, or the cloud (Amazon Web Services, Google Cloud Platform, Microsoft Azure). [@Wright_Computing].

Different researchers, especially those located at different institutions, those at different career stages, and those who have different research focuses have different computing environments (potentially differences in operating systems, software, software versions, etc.). Beyond differences in operating systems, different institutions and the IT departments that manage the institution's computers may have specific setups and guidelines shaping the available software and computing environment for computers at the institution. Individuals with different needs will install and update software differentially. Even team members within the same lab will have different computing environments. Overtime, a single researcher's computer will evolve as well. Software and operating systems will be updated (sometimes automatically rather than manually) and new software, tools, or applications will be installed. Basically, computers are like snowflakes -- they are unique when compared to other computers. They make have similarities, but overall computers are their own unique entity.

```{r fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = 'Within a team or collaborative environment, each computer is unique – like snowflakes.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f6b98e19f5_0_1#slide=id.g3f6b98e19f5_0_1")
```

The differences between computers can be problematic for performing and sharing reproducible research. 


```{r fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = 'Containers provide a consistent computing environment for everyone on the team, isolated from the unique setups.'}
ottrpal::include_slide('https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f6b98e19f5_0_49#slide=id.g3f6b98e19f5_0_49')
```

***
