# **Introduction to Containers**

***

Before we start to address your project with Collab O. Rator, we will provide an introduction to {{< glossary "Containerization" display="containerization" >}} with {{< glossary "Docker" >}}. In this section, you will learn more about what containers are, why they are useful, and when they may be unnecessary. In additional you will learn about different software tools or technologies that support containerization, when certain technologies are more appropriate to use, as well as how to find, share, and work with containers. Towards the end of this section we will return to the {{< glossary "Microbiome" display="microbiome" >}}-{{< glossary "Metabolome" display="metabolome" >}} data collection project.

## What is a container?

***

A [container is an active computing environment](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-a-container/) that is isolated from the rest of the user's system. Containers are the operational ending point of the containerization process. The containerization process also involves Dockerfiles and images (upstream of containers) which define and instantiate or produce containers. 

Images are a frozen snapshot of what the computing environment will be in a container or what will be in the running container, while a container is the isolated, specified computing environment that you can work in. This is like the Mary Poppins scene where Mary Poppins, Bert, and the kids are looking at chalk drawings (images) and jump into one which has its own isolated environment (container). 

```{r, echo = FALSE, fig.alt = "Images are a frozen snapshot of what the computing environment will be in a container or what will be in the running container, while a container is the isolated, specified computing environment that you can work in. This is like the Mary Poppins scene where Mary Poppins, Bert, and the kids are looking at chalk drawings (images) and jump into one which has its own isolated environment (container)."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f020ee26d5_2_0#slide=id.g3f020ee26d5_2_0")
```

Containers and the containerization process are supported by different tools such as Docker, Podman, Apptainer, and Singularity. Docker is one of the industry-standard and most popular of the tools and primarily will be the one discussed and utilized within this case study. However, Docker is not always the appropriate tool to use in every context.

We will discuss the containerization process that leads to containers, starting at the beginning:

### Dockerfiles

Dockerfiles are the starting point in the containerization process. Dockerfiles are text-based scripts that contain the commands that will be used to provision the eventual container computing environment. These commands will be evaluated sequentially one after the other when the Dockerfile is "built" to produce an image. In this way, Dockerfiles are a blueprint, specifying what the eventual container will have available in its computing environment for the user. Dockerfiles are small and relatively easy to store and read. Because they are text-based, they can be version-controlled. The sorts of [commands that will be in Dockerfiles](https://docs.docker.com/get-started/docker-concepts/building-images/writing-a-dockerfile/) include specifying a base image, which dependencies or software packages should be installed, a file structure to set up, defining persistent environment variables, etc. 

```{r, echo = FALSE, fig.alt = "Dockerfiles are the starting point in the containerization process "}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f3ca37d253_0_0#slide=id.g3f3ca37d253_0_0")
```

For other tools (e.g., Singularity or Apptainer) the equivalent of these files are instead called definition files.

### Building

"Build" is the verb that is used to move from a Dockerfile to an image within the containerization process. Dockerfiles are built to produce images. Images are built in a way that the resulting containers will be able to run on specific platforms or architectures (e.g., hardware + operating system: linux or windows). [Multi-platform builds](https://docs.docker.com/build/building/multi-platform/) allow for even greater flexibility in who can use a container no matter what hardware + operating system they have. For example, newer Macs have a different processor chip (Apple silicon) than older macs (Intel) and may have more difficulty building, pulling, or running older images. [See this description about known issues](https://docs.docker.com/desktop/troubleshoot-and-support/troubleshoot/known-issues/) for ways around this. 

[Best practices for writing a `Dockerfile` to optimize the building process](https://docs.docker.com/build/building/best-practices/#run) are provided by Docker documentation. Many of these will be discussed later in the Data Exploration section.

### Images

[Images are layered, frozen snapshots](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-an-image/) that contain all of the necessary files, configurations, package libraries, etc that are necessary to produce the active running container. An image is not an active environment, but rather an intermediate step in the containerization process. Each layer is related to a command from the Dockerfile: adding or modifying to what the container will contain. Once images are created they are immutable (e.g., cannot be changed or modified). If you need something to be changed, you would have to create a new image. You may want to use the same image name but a different "tag" if it is a small change or updated version of an old/no longer necessary image. Common tags include "main", "latest", "dev", "v#", etc. Because images are immutable and layered, you can use images as bases for other images, starting with an existing image and adding a package or set of packages on top of it (e.g., the [base image](https://docs.docker.com/build/building/base-images/) that is specified in the Dockerfile). Images are much larger than Dockerfiles (often Gigabytes (GB) in size) because they contain the actual software libraries. Also images are not as readable as a Dockerfile, but images are the main way that containers are shared. 

Because images are the main way that containers are shared and they tend to be large already, it's an important best practice when working with containers to keep a container "minimal" -- being careful to include only what you need. Larger images require more storage space, may require more memory, take longer to "build" and "pull" (slowing down a project), and pose greater security vulnerabilities because they include more. Smaller images are also easier to debug and maintain.

:::{.emphasis_block}
<details><summary>Picking Base Images and keeping an image "minimal"</summary>

Picking the right base image when writing a Dockerfile requires a balance: finding one that has the packages/ecosystem you need while not including too much extraneous stuff you don't need (e.g., compatibility vs size).

When defining what your container should look like, there are two important considerations as [discussed in the ITN Containers for Scientists course](https://hutchdatascience.org/Conatainers_for_Scientists/05-modifying-containers.html#activity-instructions):

1. Start with a small base image that provides the ecosystem you need (e.g., Python or R) and add only the fewest number of packages that you need for your specific project. 
2. If a project has a lot of steps or would need a very large image, you could split the project into smaller units and have a docker image for each segment of the project. This is especially useful if the project is utilizing workflows (e.g., WDL, Snakemake) where different containers can be specified for different steps.

Some base images on Docker Hub (such as Python) provide [`slim`](https://hub.docker.com/_/python#pythonversion-slim) or [`alpine`](https://hub.docker.com/_/python#pythonversion-alpine) options that are scaled down in what they include and are helpful when size is a major concern. However, they may also cause compatibility issues if you need to install and use some additional packages. 

</details>
:::

```{r, echo = FALSE, fig.alt = "Images are layered and built by processing the sequential commands in a Dockerfile, stacking each layer on the previous layer according to the commands in the Dockerfile."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f01a9d8250_0_17#slide=id.g3f01a9d8250_0_17")
```

### Registries

Images can be stored within your local file system but you might want to free up space (due to their large size) or share them with team members, collaborators, another machine you use, or publicly because you are publishing a study. This is where [registries](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-a-registry/) are useful. A registry is a centralized location for storing and sharing container images. Docker Hub is a common, but not the only, registry used to share Docker images. Registries like Docker Hub are useful for finding base images or full images that can be used as is for your particular use case.

```{r, echo = FALSE, fig.alt = "Images are stored in registries and can be pushed or pulled from a registry. The images that are pushed to a registry are built from a Dockerfile. The images that are pulled from the registry will match the images that were pushed. If you pull an image and later use the original Dockerfile to build an image, these could be the same image or they might be different if undefined package or dependency versions have changed."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3fb50cd5a7c_0_323#slide=id.g3fb50cd5a7c_0_323")
```

Images can be "pushed" or "pulled" from a registry. If an image has multiple tags or are built for multiple architectures, the related images (with different tags or built for different architectures) will be stored in a [repository within the registry](https://docs.docker.com/get-started/docker-concepts/the-basics/what-is-a-registry/#registry-vs-repository).  If an image contains proprietary information, it likely shouldn't be shared in a public registry like Docker Hub.

:::{.column-margin}
Add push and pull to the glossary as well as a dictionary box.
:::

:::{.reproducibility_box}
::::{.reproducibility_box_header}
Reproducibility considerations
::::
::::{.inner_block}
An image that is pulled from a registry will match the image that was pushed to the registry. If after some time (after pushing the image to the registry) an image is built from the same `Dockerfile` that was used to build the original image (the one that was pushed to the registry), these images may no longer match. Unless versions are explicitly controlled within the `Dockerfile`, versions of dependencies or packages may have changes or may no longer be available. [See this blog for more information](https://oneuptime.com/blog/post/2026-02-08-how-to-pin-package-versions-in-dockerfiles-for-reproducible-builds/view). Also, you may be using different options or parameters to build the Docker image than were originally used. Small changes due to differing build options and time passing can impact how reproducible a container actually is in practice and is a reason why Images are typically stored in registries to share rather than sharing a `Dockerfile` (which may or may not version pin all software). 
::::
:::

### Running

"Run" is the verb that is used to move from an image to a container within the containerization process. Images are run to create or launch a container -- the active, usable computing environment. Once a container is running, files from your system cannot be added to the container because the container is isolated. You can likely use [`wget`](https://www.geeksforgeeks.org/linux-unix/wget-command-in-linux-unix/) to fetch files from the internet, but those files will not be available to others using the same image to run another version of the container unless they fetch them as well. 

### Mounting

When you "run" an image to create a container you can mount a volume of your local files so that files from your system are available to the container. These files may be a analysis scripts, data files, package that you are writing/developing. Any changes made to those files will still be available to you once the container is no longer running.

```{r, echo = FALSE, fig.alt = "Volumes have to be mounted when running an image to have access to outside files such as scripts or data."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f613e4255f_0_0#slide=id.g3f613e4255f_0_0")
```

If your institution has strict security configurations set on your machine (possibly due to sensitive data), you may not be able to mount volumes when running a Docker container. This is because Docker as a tool requires root-level permissions and this goes against the [principle of least privilege](https://hutchdatascience.org/Ethical_Data_Handling_for_Cancer_Research/data-security.html#the-principal-of-least-privilege), an important consideration for data management. 

:::{.emphasis_block}
<details><summary>Alternatives for machines with strict security configurations</summary>

* [Podman](https://podman.io/) may be able to be used and has many similarities to Docker while more closely aligning with the principle of least privilege.
* [Docker can work with singularity](https://docs.sylabs.io/guides/2.6/user-guide/singularity_and_docker.html), enabling you to still use Docker and Docker containers without needing root access.

</details>
:::

### Vocabulary Summary

::: {.emphasis_block}
In summary, Dockerfiles are text-based scripts that contain commands defining how an image will be built. Dockerfiles serve as a blueprint or the starting point of the containerization process. Dockerfiles are built to produce images. Images are a frozen, layered snapshot of what the container will be, containing the files for the system libraries. Images are much larger than Dockerfiles and less readable. They are the most common way to share a container and are often stored in registries (like Docker Hub). Images are run to create containers and file systems or volumes may be mounted during the run process so that isolated containers have access to scripts or data and any updates to those are available outside of the container as well.
:::


## Containerization Analogies

***

::: {.emphasis_block}

<details><summary><b>Containerization Analogies</b></summary>

::: {.panel-tabset}

### Kitchen and cooking analogy

The containerization process can be thought of like working in a kitchen for dinner service or a holiday meal. 

```{r fig.align = 'center', out.width = '100%', echo = FALSE, fig.alt = 'Diagram showing how working in a kitchen can be an analogy for the containerization process. Dockerfiles specify the recipe repertoire and kitchen organization standards while building is like the preparation process resulting in standardized resources, ready for active cooking, but not actively being used - an image. Running the image results in a container or an active kitchen where cooking is happening. It is like turning on the burner. In the running process you may mount a volume of scripts and data or in the kitchen this would be specific recipes. Finally, container clean up is removing the heat and resetting the environment by storing ingredients and cleaning up the space.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f88b06ed70_0_0#slide=id.g3f88b06ed70_0_0")
```

In the kitchen you have a recipe or a set of recipes each listing the necessary ingredients to stock as well as how much of each ingredient is needed and if there's any special preparation or considerations for an ingredient (e.g., chopping or peeling). Recipes also provide detailed instructions for the environment (desired pans, cooking temperature, time). Additionally, a kitchen has specific locations for specific tools, pans, and utensils -- perhaps there's a map or binder specifying where things are located and how they are to be cleaned and stored. All of these together specify how the kitchen should be provisioned.

The Dockerfile is like a recipe and kitchen binder specifying what dependencies, packages, and environmental variables (ingredients) are needed, in what order, and what the environment needs to look like. While a Dockerfile contains a layered, step-by-step definition of the computing environment, it doesn't contain a step-by-step definition of how a specific analysis will utilize the computing environment. 

Working in a kitchen, often the area and ingredients may be prepped before any combining, cooking, or baking actually happens. Pans, utensils, and tools are all cleaned and stored in specific, known locations. Ingredients that need peeled, chopped, or some other special preparation are prepared in those ways. This preparation allows for a more efficient or easier combination of ingredients within the cooking process. Everything is prepared (according to kitchen station expectations and the instructions for ingredients in the recipe), just sitting at the ready, but not actively being used. This is termed "mise en place" (pronunciation: /ˌmiz ɑn ˈplɑs/ or meez ahn PLAHSS).

The "build" process (Dockerfile --> image) is the preparation of the environment and ingredients: the cleaning, stocking, chopping, dicing, peeling, measuring, etc. which leads to the docker image where everything is in a static, read-only, prepared state. The docker image is like "mise en place".

The step-by-step instructions for specific dishes, and their utilization of specific ingredients may be needed as well (analysis scripts). Perhaps unusual ingredients (data) for special dishes are needed. These are not part of the building process, but instead are brought into the kitchen (container) once everything else is ready. The recipes, kitchen station and prep guidelines (Dockerfile), and prepared environment, and ingredients (image) provide a static and standardized setup -- one that can be used by any set of cooks or chefs. The mounted volume within a running container brings in the flexibility to use the static and standardized setup for your specific needs in your specific way.

Once the kitchen is ready, they fire up a stove top, heat an oven, or combine ingredients that will work together in a bowl. With the addition of heat or mixing of ingredients, the cooking process itself begins. Everything that was defined in the recipe or kitchen station instructions is at the kitchen's disposal. It's too late to go get different ingredients or supplies.

The "run" process (image --> container) is like supplying the heat -- creating the environment where the cooking and combining of ingredients can happen -- the container. In the running container or environment, each of the ingredients (packages) are at your disposal, ready to use. Your analysis scripts and specific data will be incorporated here by mounting a volume during the "run" process. 

Within the running container, you use the packages together with any data or analysis scripts to perform an analysis or provision a tool. Everything defined in the Dockerfile is at your disposal to use in the running container. And the running container is isolated from the rest of your computer -- so only what is defined in the Dockerfile is available unless you've mounted a volume to bring in specific data and scripts.

The eventual meals produced by the kitchen are like the finished product: a tool, analysis, result, etc. that can be shared. And finally, as with any kitchen endeavour -- there has to be a clean up. Perhaps a container is paused or stopped (e.g., the heat is turned off but the pan is still available on the stove). Then eventually the container is removed -- (e.g., dishes are cleared, cleaned, and put away). Images may also be removed to save space -- like clearing out or storing unused ingredients. 

In a kitchen, you may make multiple courses for a meal: an appetizer, the main course, and dessert. Each of these have their own process: own station in a professional kitchen, own set of recipes, own prep needs, and own tool needs. However, the end product is the whole meal. In a similar way, an analysis may have multiple parts each supported by its own container, all working together to produce the overall end product.

### Hotel analogy

The containerization process can be thought of like hotels. 

```{r eval = FALSE, fig.align = 'center', out.width = '100%', echo = FALSE, fig.alt = 'Diagram showing how the containerization process can be thought of as hotels.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f8b4188434_0_21#slide=id.g3f8b4188434_0_21")
```

* Dockerfile: The blueprint for the room plus the instructions for cleaning/resetting it in between guests.
* Build: Cleaning/resetting the room in between guests.
* Image: The room cleaned and prepped for check-in. It's identical every time you look at it.
* Run: Check-in/ handing the room key to the guest.
* Container: The active occupied room during a specific guest's stay.
* Note each room is isolated from the others and the amenities that can be shared by multiple rooms can be thought of like a mounted volume.
* Cleanup steps: checking out.
* Different hotel chains will look different and may provide different amenities (base images), but they will all have similarities in providing a bed, bathroom, etc. In a similar way containers will have different packages and you bring in different date or scripts when mounting a volume, but they all are similar in that they provide an isolated environment.


### Lego analogy

The containerization process can be thought of like building with LEGOs.

```{r eval = FALSE, fig.align = 'center', out.width = '100%', echo = FALSE, fig.alt = 'Diagram showing how the containerization process can be thought of as building legos.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f8b4188434_0_63#slide=id.g3f8b4188434_0_63")
```

* Dockerfile: Instruction booklet that comes with a set
* Build: LEGO assembly line: gathering the necessary pieces, packing them into sealed bags, putting them and the manual (what's necessary for the set) in a sealed box
* Registry: Lego.com or a LEGO store
* Image: The sealed box for the set with pieces in bags and booklet included.
* Run: Opening the box and dumping the pieces on the table.
* Container: Actively putting the set together and playing with it.
* Different sets will have different things available the same way that different containers will have different things available.
* Mounting a volume: bringing LEGOs that you have outside of the set into the act of putting the set together to modify it or make it your own. 

:::

</details>

:::

:::{.column-margin}
Will embed a survey to ask learners to suggest other analogies.
:::

## Why containers?

***

As previously described, containers are useful for standardizing the computing environment within a team or between different machines (e.g., local laptop and institutional HPC or cloud computing environment).

Containers are not always necessary. The complexity of setting up, sharing, and documenting containers may not be worth what you gain in regards to reproducibility. In such cases reporting software used and their versions is often enough. Such cases include: 

* If you're working on an analysis without specialized software. Specialized software is software that may be hard to install because of dependencies or software that you wouldn't expect many people in your field to work with.
* If you're **not** working within a large collaborative team.
* If your analysis is tied to a single machine or static compute environment, and won't need to move between systems or machines.
* If the computing platform already supports analyses through the use of containers (e.g., The NHGRI AnVIL Platform), then you may not need to set up your own container if theirs already provides the software you need.

In cases where you don't need a container, consider printing out [Session Info](https://sessioninfo.r-lib.org/) or setting up an [`renv`](https://rstudio.github.io/renv/) to more transparently report your reproducible environments for your analyses.

### When is a container especially useful?

Containers are especially useful in the following cases:

* If you need to move an analysis to another machine or environment like an HPC or the cloud.
* If a collaborative team is working on a large analysis within a workflow (e.g., WDL, Snakemake) or in different environments.
* If a collaborative team is having difficulty replicating results within the team.
* If a team wants to both replicate and extend an analysis.
* If there are many dependencies that could cause nightmares.

```{r, fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = ''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f020ee26d5_4_12#slide=id.g3f020ee26d5_4_12")
```

If you need to move an analysis to another machine or environment like an HPC, it is important to know which containerization tool is the most appropriate. Many institutional HPCs do not allow Docker use because of its required root-level execution. This would give users higher permissions than they generally need, acting counter to the principle of least privilege, and creating unnecessary security vulnerabilities. Singularity and Apptainer are more commonly supported / used by institutional HPCs. When working on the cloud, Kubernetes is a common tool to orchestrate container deployment and manage large-scale Docker container infrastructure.

### What are general use cases for containers?

There are several different (downstream) use cases for containers: 

* Supporting an analysis: Due to the clear and consistent definition and packaging of what is in a container, the ability to share it among a team or across platforms (portability), and the ease with which it can be integrated into workflows, containers are useful for supporting a computational analysis. Containers are especially useful in the team environment for supporting an analysis when it comes to debugging and reproducibility. Team members may find that analyses have different results if they do not have standardized software versions and computing environments.
* Sandboxed software evaluation: Due to the isolation of the container environment, containers provide an isolated sandbox where software can be tested or evaluated without risking the security of the rest of the system or being impacted by other software [@Tommaso_Palumbo_Chatzou_Prieto_Heuer_Notredame_2015].
* Deploying an app: Containers support deploying applications because of their consistency, portability, and isolation. The application's code and dependencies are all packaged together and will run the same way every time no matter what servers or cloud services the app is on (portability). If there's an error, bug, or crash it won't impact other apps on the server because of the app's isolated environment.

When trying to decide if containers are useful for you and your work, consider which of these use cases you need. If your use case is supporting an analysis and you don't need the ability to share and integrate with workflows, a container may be unnecessary. The first of these use cases (supporting an analysis) is the context for this case study.

### What are the first steps for working with a container?

If a container is necessary for you work, the first steps largely depend on how the container will be used and what stage the project is in. For a project that is starting from the beginning, you may need to develop the container from the beginning by defining a Dockerfile. This would include choosing a base image and specifying dependencies and the actual packages of interest. Alternatively, you may be able to find a container in a registry that meets your needs already because it was referenced in a published paper. As the project evolves, you may have to further modify the container so that it contains more or different packages than you expected. Be sure to remove any unnecessary packages that you didn't actually need. For projects that are more mature, a container may already be largely defined / available. In such cases a team member or collaborator may have to give you access instructions.

The last of these scenarios (already specified container) is the context for this case study.

## How do you work with a container?

***

Using Docker through Docker Desktop vs through a Command Line Interface (CLI) has some trade-offs. Beginners may find it easier to work within Docker Desktop because of its Graphical User Interface (GUI): users will be able to see lists of images and containers, buttons to control container actions (such as running or stopping), lists of files within mounted volumes, do everything within the GUI without having to switch between windows ([Explore what the Docker Desktop GUI has.](https://docs.docker.com/desktop/use-desktop/)). Docker Desktop also has [extensive documentation](https://docs.docker.com/desktop/) about working with the GUI. However, [Docker Desktop does not provide tab completion by default](https://docs.docker.com/desktop/settings-and-maintenance/settings/), and once you enable it, it doesn't work quite in the way you would expect it to while interacting with and running files in the "Exec" tab. Additionally, Docker Desktop can be a (manageable) resource drain on a system, utilizing a large amount of memory and battery.

Depending on how a container is defined / setup, it may install an interactive Integrated Development Environment (IDE), such that you can work within an IDE (e.g., RStudio/Positron, Visual Studio Code, Jupyter Notebooks, etc.) within a running container. Docker can also be used with workflows (e.g., WDL, Snakemake, GitHub Actions). Using containers with GitHub Actions or within the cloud moves where you run the downstream interaction from your local machine to a remote location.

Using Docker through Docker Desktop or through a CLI (locally) will be the context of this case study.


## Summary

***

::: {.emphasis_block}

Beyond the decision of whether containers are the right choice for you and your downstream use/work, working within the containerization process requires decisions: what you need available in your container; what can be shared and where; what files you'll need to mount (if any) when you run the container; whether you're running the container locally or remotely.

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f98af6272b_0_12#slide=id.g3f98af6272b_0_12")
```

:::

***

