# **Context**
***

## What is the metabolome?

The metabolome is a collection of metabolites or low-molecular weight organic (containing carbon-hydrogen bonds) molecules found in a specific sample or organism. Large organic biomolecules in the cell are primarily categorized among four groups: nucleic acids (DNA or RNA), proteins, carbohydrates, and lipids [@Cooper_2000]. These categories represent large biomolecules (macromolecules) broadly responsible for cellular structure, energy storage, enzymatic catalysis, and genetic information storage and transmission. Metabolites may be by-products from, precursors, intermediates, or building blocks of the various macromolecules. They also include vitamins, organic acids, and other small organic molecules that don't neatly relate to the larger classes of biomolecules [@Karu_Deng_Slae_Guo_Sajed_Huynh_Wine_Wishart_2018]. 

Metabolites may be endogenous or exogenous depending on their origin. Endogenous metabolites originate from the organism or source of the sample while exogenous metabolites originate from outside of the organism / source of the sample. Metabolites may be classified as either primary or secondary metabolites where primary metabolites are required for the life-cycle: growth, development, and reproduction; secondary metabolites are instead used for non-life cycle tasks such as communication/signaling or defense. Additionally, metabolites may be water-soluble or water-insoluble, each requiring their own sample extraction techniques [@Lu_Su_Klein_Lewis_Fiehn_Rabinowitz_2017]. 

Metabolites are detected, identified, and quantified using nuclear magnetic resonance (NMR), enzyme assays, or by pairing compound separation through Gas Chromatography (GC) or Liquid Chromatography (LC) with mass spectrometry (mass spec or MS), and working with the data from those experiments in downstream data processing [@Lu_Su_Klein_Lewis_Fiehn_Rabinowitz_2017]. A typical output of NMR and MS experiments are visual graphs called spectra.  

::: {.column-margin}
Add a figure with an example of a spectrum output
:::

<details><summary>Resources for more details on NMR and mass spec</summary>

**NMR**

For more details on how NMR works, visit this [resource from the University of Wyoming about the basics of NMR spectroscopy](https://www.uwyo.edu/chemistry/instrumentation/NMR/Introduction.html).

Also consider @Marion_2013 for "An Introduction to Biological NMR Spectroscopy" from the Journal of Molecular & Cellular Proteomics.

**Mass Spec**

For more details on how mass spectrometry works, visit this [resource from the Broad Institute explaining what mass spec is](https://www.broadinstitute.org/technology-areas/what-mass-spectrometry), this [resource from Scripps Research about the basics as well a more thorough discussion on the relevant technologies](https://masspec.scripps.edu/learn/ms/) or this [resource from Michigan State University about the basics with an associated practice problem set](https://www2.chemistry.msu.edu/faculty/reusch/virttxtjml/spectrpy/massspec/masspec1.htm).

Additional resources include @Garg_Zubair_2026 and @Son_etal_2024.

</details>

In human fecal samples, a mixture of metabolites would be expected -- both endogenous and exogenous -- due to physiology (endogenous), dietary intake (exogenous), and gut microbiota metabolism (considered exogenous due to the secondary metabolism of endogenous metabolites) [@Liu_Tan_Cheng_Zhang_Feng_Peng_2022]. Classes of expected metabolites that will be observed include fatty acids (a lipid building block), bile acid (steroid derivatives synthesized from cholesterol), amino acids and protein derivatives, intermediates of carbohydrate metabolism, and undigested dietary sugars [@Karu_Deng_Slae_Guo_Sajed_Huynh_Wine_Wishart_2018]. 

Historically, the "low-molecular weight" part of the definition of metabolites has been thresholded with a cutoff for molecules: <1500 Daltons in molecular weight. A recent article [@Moreno-Ulloa_2026] proposed a re-definition of metabolites that no longer restricts their identity based on size. However, this definition does not seem to be a widely accepted by the field at this time.

To aide the downstream data processing (because of the vast number of metabolites that have been discovered and studied, their possible locations, and their many associations with different signaling, gene-regulatory, and metabolic pathways or disease implications), dedicated reference databases with known metabolites and related info were developed and are maintained by the scientific community. Some of these databases provide specialized metabolite information across a range of organisms where these metabolites would be found such as spectra from NMR or MS experiments (e.g., MassBank of North America or MoNA) [@MassBank], names, structural, or chemical information about a specific class of metabolite (e.g., LIPID MAPS) [@Conroy_etal_2024], or metabolic pathways/networks (e.g., Kyoto Encyclopedia of Genes and Genomes or KEGG) [@Kanehisa_Goto_2000; @Kanehisa_etal_2025], while others provide a more comprehensive, but organism-specific databases like the Human Metabolome Database (HMDB) [@Wishart_etal_2009].

From the 2009 paper introducing HMDB [@Wishart_etal_2009], the database is described in the following way:

> HMDB is ... [an] organism-specific metabolomics database .... It contains spectroscopic, quantitative, analytic and molecular-scale information about human metabolites, their associated enzymes or transporters, their abundance and disease-related properties.


The metabolome of an individual may change over time. Therefore, some research looks at the metabolome of the same research participants over a course of time. This is called a **longitudinal** study. Research participants within the same study (whether it is longitudinal or not) are termed a **cohort**. 

<details><summary>What are different applications of metabolomics within the scientific field></summary>

Some applications:

* Biomarker identification
* Drug discovery or development
* Clinical Toxicology
* Nutritional studies
* Quantitative phenotyping
* Interpretation and understanding of complex biological processes

</details>

## How are the microbiome and the metabolome related?

The microbiome is discussed in more [detail within the Version Control Case Study](https://www.opencasestudies.org/ocs-bio-version-control/#what-is-a-microbiome), but briefly, the microbiome is the set or collection of microorganisms (e.g., bacteria, viruses, fungi, etc.) that live in a particular environment such as the human gut. These microorganisms can impact metabolism and have been found to be associated with greater human health.

Sequencing the microbiome can help us to identify what bacteria and other microorganisms are present in the human gut. Different proportions of different organisms can have  an impact on human health. However, to actually see the activity of or impact caused by the presence of those microorganisms, other experiments such as the identification of metabolites in the samples are helpful. Considering connected data like this in an analysis is termed an **integrative** analysis because it is integrating multiple sources of data.

Note that within the context of the microbiome, a "secondary metabolite" has an additional meaning besides ones that are used for non-life cycle tasks. Secondary metabolites may also refer to metabolites that are from the primary "host" system but are then re-metabolized by the microbiome [@Liu_Tan_Cheng_Zhang_Feng_Peng_2022]. 

Because microbiome and metabolome data can be integrated together for a greater depth of scientific insight and conclusions, often studies will "pair" microbiome and metabolome data, meaning that for each sample in the study, the researchers will perform experiments to survey both the microbiome and the metabolome of that sample [@]. The data is then termed **paired** data. 

## What is a curated data resource?

@Muller_Algavi_Borenstein_2022 introduced and explored a **curated data resource**. Simply, this means a collection of datasets that have been gathered or curated by a group for secondary use within the scientific field. This often goes further than just collecting references of different studies who have made their data available and includes reprocessing the datasets from those studies to make sure that they have been handled in the same way and can be used together in future work. In this case, the researchers from @Muller_Algavi_Borenstein_2022, after identifying studies to include in the data resource had to 

1. Download the data associated with each study.
2. Link the paired data (microbiome and metabolome) for each sample within each study.
3. Jointly reprocess all raw data from across the studies
4. Annotate microbiome taxons and metabolite molecules according to relevant databases (e.g., KEGG and HMDB)

Recall that the set of samples from a single study is typically referred to as being from a cohort. In a curated data resource, the different cohorts are pooled together.

:::{.column-margin}
Add a figure that shows this idea of cohorts being pooled together.
:::

## What are possible uses of a data resource?

Curated data resources are useful for a variety of tasks such as:

* Working with more data for greater statistical power
* Perform a meta-analysis to compare results and associations among datasets (See the [Version Control case study for an expanded discussion on meta-analyses](https://www.opencasestudies.org/ocs-bio-version-control/#what-is-a-meta-analysis).)
* Benchmarking tools
* Comparing findings from an analysis to a different dataset (e.g., evaluating the replicability and generalizability of the research) [@Schloss_2018]


## The need for containerization

Data analysis, especially analysis of large data, relies on computers and the software tools installed on those computers. Depending on the computational needs of an analysis or a researcher's typical work, that computer or work station may be a personal computer / laptop (Operating system: Microsoft Windows, Apple MacOS, Google ChromeOS, or Linux), a desktop, a Graphics Processing Unit (GPU) server, shared computing resources such as institutional high performance computing (HPC) clusters or supercomputers, or the cloud (Amazon Web Services, Google Cloud Platform, Microsoft Azure). [@Wright_Computing].

Different researchers, especially those located at different institutions, those at different career stages, and those who have different research focuses have different computing environments (potentially differences in operating systems, software, software versions, etc.). Beyond differences in operating systems, different institutions and the IT departments that manage the institution's computers may have specific setups and guidelines shaping the available software and computing environment for computers at the institution. Individuals with different needs will install and update software differentially. Even team members within the same lab will have different computing environments. Overtime, a single researcher's computer will evolve as well. Software and operating systems will be updated (sometimes automatically rather than manually) and new software, tools, or applications will be installed. Basically, computers are like snowflakes -- they are unique when compared to other computers. They make have similarities, but overall computers are their own unique entity.

```{r fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = 'Within a team or collaborative environment, each computer is unique – like snowflakes.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f6b98e19f5_0_1#slide=id.g3f6b98e19f5_0_1")
```

The differences between computers can be problematic for performing and sharing reproducible research [@]. Several examples exist in the literature where different versions of software lead to different results when analyzing the same data [@].

:::{.reproducibility_box}
::::{.reproducibility_box_header}
Reproducibility considerations
::::
::::{.inner_block}

Reproducible research is where the same methods (code, script, implementation based on description, etc.) can be used on the same data or within the same experimental system by a new researcher to recapitulate the results or findings of the original researcher.

Note that just because a research result is reproducible doesn't mean that it's correct or valid. Research can be reproducibly wrong. However, reproducibility is always an important step towards useful, valid, and generalizable scientific discovery.

<details><summary>What is reproducible research within the greater context of scientific discovery?</summary>

**Repeatability**: Same researcher using the original data and process to repeat the same result.

**Reproducibility**: New researcher using the original data and process to repeat the same result.

**Replicability**: New researcher using new data with the original process to repeat the same result.

**Extensibility**: New researcher using the original data and process but expanding the method to repeat the same result but also produce an extension of the result.

```{r fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = 'Repeatability is the same researcher repeating the same result using the same data and methods. Reproducibility is another researcher repeating the same result using the same data and methods. Replicability is a new researcher using new data but the same process and repeating the result. Extensibility is using the same data with the original method, but also expanding it for an extended result.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f020ee26d5_4_12#slide=id.g3f020ee26d5_4_12")
```

As described in @Schloss_2018, preferably a research result can go beyond replicability and also be robust and generalizable. 

**Generalizability**: Using different methods and different data but producing the same overall result.

**Robustness**: Using different methods but the same data and producing the same overall result.

</details>

::::
:::


One way people have tried to get around this challenge is reporting the exact versions of all software that were used. This may at times be sufficient, but other times more precision is necessary. Even if given a list the exact versions of all software and packages or libraries that were previously used, a researcher would need to install any new software and update or downgrade any software they've previously installed. Some software requires special dependencies that may not be known to be required until attempting to install a specific package or library and these dependencies may be especially difficult to install due to conflicting, missing, or mismatched version requirements among other software that is already installed. Teams may have a shared computing resource (like an institutional HPC), but even in that case different users have different software and configurations. Geographically dispersed teams can't realistically mail a laptop to each other to have a standardized computing environment.  

Therefore, containers are often the solution used to standardize the computing environment within a team. Containers provide a standardized computing environment for everyone using it no matter what the software landscape of their personal laptop or workstation looks like.


```{r fig.align = 'center', out.width = "100%", echo = FALSE, fig.alt = 'Containers provide a consistent computing environment for everyone on the team, isolated from the unique setups.'}
ottrpal::include_slide('https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f6b98e19f5_0_49#slide=id.g3f6b98e19f5_0_49')
```

***
