# **Data Exploration**
***

## Container

We'll utilize this section to explore the Dockerfile specifically answering "What's in a container?"

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# 1. Fetch the content
url <- "https://githubusercontent.com"
content <- httr::content(GET(url), as = "text", encoding = "UTF-8")

# 2. Build the Markdown block manually as a single text string
markdown_block <- paste0(
  "```dockerfile filename=\"Dockerfile\"\n",
  content,
  "\n```"
)

# 3. Use knitr's native function to safely push it out as raw markdown
knitr::asis_output(markdown_block)
```

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f01a9d8250_0_17#slide=id.g3f01a9d8250_0_17")
```

* Notice there are layers
* Point out dependencies (Side quest: how do we know what dependencies we need?)
* The software or packages of major interest
* Keeping it modular or small and having multiple containers for the same project/different substeps (note that in this case study we're using the data visualization container the team is using...it'll have some visualization packages that are reasonably being used by the team for visualizations, but not the visualization steps within this case study)

We'll also utilize this section to explore what's not in a container! -- Data! (noting that yes we're storing some intermediate data in the container while we're using it, but we won't include that in a Dockerfile or an image / container we share with others)

We can even explore the file structure of the running container in Docker Desktop a bit and connect how those directories are related to what's in the Dockerfile.

Side quest: Discuss mounting data/scripts in this section (as an alternative to editing the files in the file tab) and say that we'll be doing that later on in the Data Analysis section (though this isn't always possible depending on security settings on a system)

## Application to data

Using the `tabulapdf` package to extract the data table from the research study PDF, we defined a Data Frame (`raw_table`) that will need some wrangling to make it more accurately reflect Table 1 from @Muller_Algavi_Borenstein_2022. Rather than having 14 rows and 7 columns of data as we expected, we observe a Data Frame with 25 rows of data. And while our Data Frame does have 7 columns, one of them (`X4`) is all NAs.

```{r raw_table_visualized, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = "The extracted data has 25 rows instead of the expected 14. While it has 7 columns as expected, one of those columns does not have data."}
ottrpal::include_slide('https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f96ddc37a7_0_131#slide=id.g3f96ddc37a7_0_131')
```

We would have expected the cohort descriptions to be in that third column and the number of paired samples to be in the fourth column. By more closely inspecting the data we have extracted, we instead see that `X3` contains both the cohort descriptions (boxed in purple below) and the number of paired samples information (circled in teal below); the fourth column (`X4`) has all `NAs`. The number of paired samples information seems to have been appended to the end of the cohort description rather than being in its own column. In addition, we would have also expected the data to have only 14 rows, but instead we are observing 25 rows. Several rows have `NAs` in all columns except for the third column. This is because some of the longer cohort descriptions are being split across more than one row.

```{r raw_table_x3, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = "The X3 column contains both the cohort descriptions which are sometimes multi-row as well as the number of paired samples data which is appended to the end of the cohort description. Cohort descriptions are boxed in purple to show some examples and number of paired samples are circled in teal."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f96ddc37a7_0_212#slide=id.g3f96ddc37a7_0_212")
```

Notice also that none of the columns have descriptive names.

Therefore, when we wrangle the data we will need to

  * rename the columns
  * separate out the number of paired samples from the cohort descriptions
  * reunite the multiline cohort descriptions into a single line
  * drop rows where the overflow cohort descriptions but no other data are
  * drop the fourth/empty column as well as the column containing a mixture of cohort descriptions and the number of paired samples (and any unnecessary/intermediate columns we make in wrangling)
