<!--
Update the file paths as needed throughout
-->

<!--
At the end of this section, save incremental data so that users can pause and come back later
-->

# **Data Visualization**
***

If you have been following along but stopped, we could load our wrangled data like so:

```{r, eval = FALSE}
load(here::here("data", "wrangled", "wrangled_data.rda"))
```

***
<details> <summary> If you skipped the data import section click here. </summary>

An RDA file (stands for R data) of the data can be found [here](https://github.com//opencasestudies/ocs-bp-co2-emissions/tree/master/data/wrangled) or slightly more directly [here](https://raw.githubusercontent.com/opencasestudies/ocs-bp-co2-emissions/master/data/wrangled/wrangled_data.rda). Download this file and then place it in your current working directory within a subdirectory called "wrangled" within a subdirectory called "data" to use the following code. We used an RStudio project and the [`here` package](https://github.com/jennybc/here_here) to navigate to the file more easily.


```{r, eval = FALSE}
load(here::here("data", "wrangled", "wrangled_data.rda"))
```

</details>
***

We'll use this section to make two scatter plots that are combined with `patchwork`, specifically using the modified container to run the code

* labeling the points with the `dataset_name`
* coloring the points based on `longitudinal` value
* x-axis is either `hmdb_annotated_compounds` or `kegg_annotated_compounds`
* y-axis is `num_paired_samples`

***
