# **Data Analysis**
***

## Container Application

Finding the correlation between different variables in the dataset

Looking at subsets of datasets for the same variables...

```{.R filename="04_analysis_metric.R"}
{{< include resources/analysis_scripts/04_analysis_metric.R >}}
```

::: {.panel-tabset}

### Docker Desktop

Ask learners to try to run the analysis

### Docker through a CLI

Ask learners to try to run the analysis

:::

Look at the logs.


::: {.emphasis_block}

```{r}
#| echo: false
#| output: asis

# Replace with your raw GitHub URL
url <- "https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers-analysis/refs/heads/main/analysis_results/logs/analysis_logs.txt"

# Fetch lines and print verbatim as Markdown with two spaces in between lines to enforce lines breaks
file_lines <- readLines(url, warn = FALSE)
formatted_lines <- paste0(file_lines, "  ")
cat(formatted_lines, sep = "\n")
```

:::


## Container Infrastructure

In this section we can discuss that someone else (maybe the person validating the work for resubmission) can use the container to run the analysis and get the exact same results and plot that you made because of the container. Specifically, what should you share with someone so they can rerun the analysis?

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f0102bcf57_7_5#slide=id.g3f0102bcf57_7_5")
```

```{.dockerfile filename="Dockerfile"}
{{< include resources/dockerfiles/activity/Dockerfile >}}
```

Sharing the container image vs sharing the Dockerfile


```{.bash filename="run_analysis.sh"}
{{< include resources/analysis_scripts/run_analysis.sh >}}
```

:::{.reproducibility_box}
::::{.reproducibility_box_header}
Reproducibility considerations
::::
::::{.inner_block}
Having a "run all" script that specifies what scripts were used and in what order for the analysis is a best practice for reproducibility [@Wilson_etal_2017].
::::
:::


We'll also discuss how to document where the scripts are, how to mount a volume (with analysis scripts) and point out that the validator can get the same results as the learner did by running the `run_analysis.sh` script.

Add a reproducibility box about pushing the new / adjusted scripts and plots as well as the documentation to the scaffold repo that they "used as a template/cloned" if they did the earlier extra reproducibility step.



***
