# **Data Wrangling**
***

## Revisions task

If you have been following along but stopped, we could load our imported data like so within our analysis script within the running container (without having to run the original `extract_tables` code):

```{r, eval = FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```
***

<details> <summary> If you skipped the data import section click here. </summary>

KI note: We'll need to add instructions to this about mounting or importing data if someone wants to pick up at this section.

```{r, eval=FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```

</details>
***

We'll need to add this code to our `analysis.R` file using the same steps as before:

* navigate to the files tab
* right click on the `preprocessing.R` script and click edit
* copy paste the following code into the script
* click save
* navigate to the exec tab
* use `preprocessing.R` -- adjusting based on script location and learner's file location.

```{r, eval = FALSE}
library(tidyverse)

table1_extract <- raw_table %>%
    mutate(num_paired_samples = as.numeric(str_extract(X3,                       #<1>
                                                      "\\d+$"))) %>%             #<1>
    mutate(X3Text = str_remove(X3,                                               #<2>
                               "\\d+$")) %>%                                     #<2>
    mutate(cohort_description = if_else(lead(is.na(X1), default = FALSE),        #<3>
                                        paste0(X3Text,                           #<4>
                                              lead(X3Text)),                     #<4>
                                        X3Text)                                  #<5>
           ) %>%
    filter(!is.na(X1)) %>%                                                       #<6>
    select(!c(X3, X4, X3Text)) %>%                                               #<7>
    rename(c("dataset_name" = "X1",                                              #<8>
             "ref" = "X2",                                                       #<8>
             "longitudinal" = "X5",                                              #<8>
             "hmdb_annotated_compounds" = "X6",                                  #<8>
             "kegg_annotated_compounds" = "X7"))                                 #<8>

```

1. Extracting the numbers (which represent the number of paired samples) at the end of the row to put them in their own column
2. Making a version of the text column that has those numbers of paired samples removed
3. Setting up an if else that evaluates whether the next value/row in the dataset name column is an NA value. This utilizes the `lead` function to look for the next value. In the case of the last row, we're using the `default` argument to pad the data with a `FALSE`
4. If the `lead` function returns that the next value is NA (`TRUE`), it will paste the current cohort text with the next row's text
5. Otherwise if the `lead` function returns that the next value is not NA (`FALSE`) (or is the last value/row), it will return the current cohort text only.
6. This filters to remove the rows where the dataset name is NA because we've already appended the important information from those rows
7. This removes the original X3 column (that had the cohort description plus number of paired samples), the X4 column which was all NAs, and the X3Text column which has incomplete cohort descriptions
8. Rename the columns so they are descriptive

Advanced side quest: Can challenge learners to come up with a more generalizable way to wrangle the data if there's more than one NA row that needs to be collapsed (e.g., some data has more than 2 lines) (use a dropdown section for this)

As a way to spot check that the container is doing something up to this point, we can use some outputs

```{r, eval = FALSE}
message(table1_extract[4, "dataset_name"])
message("Dimensions: ", paste(dim(table1_extract), collapse = " x "))
```

The Dimensions will have fewer rows (14 instead of 25) and the value of the dataset name should be `HE_INFANTS_MFGM_2019`

To allow users to skip import and wrangling we will save the data as an RDA file as well as a CSV file as this is often useful to send our data to collaborators. We will save this in a “wrangled” subdirectory of our “data” directory of our working directory.

```{r, eval = FALSE}
save(table1_extract, file = here::here("data", "wrangled",       # <1>
                                       "wrangled_data.rda"))     # <1>
readr::write_csv(table1_extract,                                 # <2>
                 file = here::here("data","wrangled",            # <2>
                                   "wrangled_data.csv"))         # <2>
```

1. saving the data as an RDA file within the wrangled data subdirectory
2. saving the data as a CSV file within the wrangled data subdirectory


Now that we've wrangled the extracted table data into a dataframe, we want to plan out the visualizations we'll make using this dataframe!

We want to make two scatter plots:

* labeling the points with the `dataset_name`
* coloring the points based on `longitudinal` value
* y-axis
  * `hmdb_annotated_compounds` for one plot
  * `kegg_annotated_compounds` for the other plot
* x-axis is `num_paired_samples`

But if we make two separate scatter plots, they will be two separate panels/files that won't be displayed next to each other unless we use design software like illustrator. But we could also use a package that allows us to combine plots, then the scatter plot will have two panels displayed next to each other. So let's "wrangle" the Dockerfile to include a package that will let us do that: `patchwork`.

## Container

As you can see from the visualization step above, these two panels are completely separate separate plots -- and we'd like them to be visible next to each other.... we can use `patchwork` for that, but we'll need to add `patchwork` to our visualization container....

Activity: Add `patchwork` to the container

* edit the Dockerfile (in your favorite text editor)
* build the Dockerfile to be an image (will need to do this on terminal)

***
