# **Data Wrangling**
***

## Container Application

If you have been following along but stopped, the scripts we are using are set up to load our imported data like so within our analysis script within the running container (without having to run the original `extract_tables` code). make sure your container is running (probably will need a tab structure for this): <- update to be more precise

```{r, eval = FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```
***

Even if you stop the container (as long as you don't delete it), you can restart it and the data will still be there as long as it was saved during wrangling

<details> <summary> If you skipped the data import section click here. </summary>

KI note: We'll need to add instructions to this about mounting or importing data if someone wants to pick up at this section. I don't think the container has `wget`, but I could set up a download that would have the file that they could then mount. Will do that and include instructions here for that.

```{r, eval=FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```

</details>
***

### Wrangling the data

::: {.panel-tabset}

We have a file that is part of the scaffold code that will wrangle the data in the ways we need.

```{.R filename="02_wrangle.R"}
{{< include resources/analysis_scripts/02_wrangle.R >}}
```



1. Rename the columns so they are descriptive
2. Extracting the numbers (which represent the number of paired samples) at the end of the cohort descriptions to put them in their own column
3. Making a intermediate column, specifically a version of the text column that has those numbers of paired samples removed
4. Setting up a `case_when` that evaluates whether the next value/row in the dataset name column is an NA value. This utilizes the `lead` function to look for the next value. In the case of the last row, we're using the `default` argument to pad the data with a `FALSE`
5. If the `lead` function returns that the next value is NA (`TRUE`), it will paste the current cohort text with the next row's text
6. Otherwise (e.g., if the `lead` function returns that the next value is not NA (`FALSE`) or is the last value/row), it will return the current cohort text only.
7. This filters to remove the rows where the dataset name is NA because we've already appended the important information from those rows to the previous rows
8. This removes the `X4` column which was all `NAs`, the original third column (that had the cohort description plus number of paired samples), and the intermediate column which we made that has incomplete cohort descriptions

Discuss what the above script is actually doing.

### Docker Desktop

how to run analysis in running container

### Docker through a CLI

how to run analysis

:::

### Verifying the wrangled data

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**What does the output from the messages in the script tell us about our wrangled data?**:

Certain lines have certain information about the data -- what are they reporting and what do we expect them to say?

1. The dimensions of the data frame
2. The number of NAs within the data frame
3. The dataset name for the fourth row of data (spot checking a value)

<details><summary>Answer</summary>

If we were to inspect the `table1_extract` Data Frame, we would observe that the data now appears to match our expectations. 

```{r}
#| echo: false
#| warning: false
#| message: false

url <- "https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers-analysis/refs/heads/main/wrangle_data/wrangled/wrangled_data.csv"

table1_extract <- read_csv(url)
table1_extract %>% print(n=14)

```

There appear to be 14 rows and 7 columns, though those columns don't follow the original order of Table 1 in @Muller_Algavi_Borenstein_2022. Since the columns have descriptive names though, it will be fine that they are not in the same order. While the columns may not be in the same order as the source data, the rows should be in the same order as the source data. Spot checking a specific value will help us to verify this.

```{r wrangled_data, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = 'The wrangled data appears to match our expectations.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f5db0d7ede_0_0#slide=id.g3f5db0d7ede_0_0")
```

The log messages report the following:

```{r}
#| echo: false
#| output: asis

# Replace with your raw GitHub URL
url <- "https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers-analysis/refs/heads/main/wrangle_results/logs/wrangling_logs.txt"

# Fetch lines and print verbatim as Markdown with two spaces in between lines to enforce lines breaks
file_lines <- readLines(url, warn = FALSE)
formatted_lines <- paste0(file_lines, "  ")
cat(formatted_lines, sep = "\n")
```

The dimensions properly report a decrease in the number of rows compared to the raw extracted data (14 instead of 25), matching our expectation for both the number of rows and the number of columns given the source data. The dataset name in the fourth row is `HE_INFANTS_MFGM_2019`, again matching our expectations when comparing to the source data.

Log message purpose | Expected | Observed
:-------------------|:---------|:--------
New dimensions      | 14 x 7   | 14 x 7
New number of NAs   | 0        | 0
Specific value check| `HE_INFANTS_MFGM_2019` | `HE_INFANTS_MFGM_2019`

</details>
::::
:::

To allow learners to skip import and wrangling in later sections, we will save the data as an RDA file as well as a CSV file as this is often useful to send our data to collaborators. We will save this in a “wrangled” subdirectory of our “data” directory of our working directory.

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**A more generalizable and robust wrangling approach**:

This wrangling approach using the `lead` function to assess if there is multiline data that needs combined assumes that there are only ever 1 - 2 rows for each cohort description. What changes to this process could be utilized that would be more generalizable or robust with data that doesn't meet this assumption? Specifically, how could this process be changed to combine an unknown number of rows of data to the previous row when only one column has data?

<details><summary>Answer</summary>
One possible approach would:

1. Use the `cumsum()` function to assign a group ID to the rows.
2. Use the `group_by()` function to group the data by that group ID.
3. With the `summarize()` function, `paste()` the cohort descriptions within each group together.
4. Either bring along the non-NA data for the other columns within the `summarize()` function, or join the datasets together using the group ID and drop unnecessary rows.

</details>
::::
:::

### Planning data visualizations

Now that we've wrangled the extracted table data into a Data Frame, we want to plan out the visualizations we'll make using this Data Frame!

Collab O. Rator's email provided the following guidance regarding the data visualizations:

>Specifically, I’d like to see scatter plots comparing the data from the final four columns for each dataset (No. samples w/ paired data, Longitudinal, No. HMDB-annotated compounds, and No. KEGG-annotated compounds).

Collab O. Rator's instructions have asked us to visualize 4 variables or dimensions of the data with scatter plots. The axes (x- and y-) will account for two of those variables, and we can adjust the points (e.g., color or shape) to represent a third variable -- likely the `longitudinal` value as this is categorical ("Yes" vs "No"). To visualize the fourth variable, we would need to add a third axis (z-) or some other means of encoding the data, but this adds to the complexity of the plot. Instead, we could make two separate panels.

Plot element | Data column
-------------|------------
x-axis       | `num_paired_samples`
y-axis       | `hmdb_annotated_compounds` or `kegg_annotated_compounds`
point color or shape | `longitudinal`
labels               | `dataset_name`

But if we make two separate scatter plots, they will be two separate plots/files that won't be displayed next to each other unless we use design software like Illustrator to combine them later. Certain R packages allow you to combine plots and display them with each other. So let's "wrangle" the Dockerfile to include a package that will let us do that: `patchwork`.

Could potentially have learners run the data visualization script that produces two separate data visualizations, but I think that's best kept for the next section?

## Container Infrastructure

As you can see from the visualization planning above, these two plots would be completely separate -- and we'd like them to be visible next to each other. A useful R package for this purpose is `patchwork`. However, our data visualization Docker container doesn't include that package.

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}

**How can we add a package to a container?**

<details><summary>Answer</summary>

There are multiple ways to add a package to a container.

If the container is running, we could install the package, and it will be usable/available, while we are using the running container (or if we pause and later restart that container). However, that package won't be available to someone else who is using the same image that we used. 

Another way to add a package to a container is to go back to the beginning definition of or recipe for that container (a Dockerfile), modify it, and then rebuild an image, and run the rebuilt image so that a new container has the additional package. 


Approach      | Pro  | Con
:-----------  | :---:|:---:
Install package in running container | Fast and we can continue our work quickly | Package not available to others using the same image we did
Modify the Dockerfile, then build, share, and run an image | All software we used is defined and available to others  | Longer containerization process before we can continue our work 


</details>

::::
:::

Wanting to make sure that we're practicing reproducibility best practices when working with containers, especially in a team environment, we're going to add a package to our container by modifying the Dockerfile, rebuilding the image, and running the container from the new image.

### Adding a package

Now we will add `patchwork` to the container -- or at least get started with the process of adding the `patchwork` package to the container.

With your favorite text editor (e.g., RStudio, Visual Studio Code, Notepad++, etc.), modify the Dockerfile to also install the `patchwork` package.

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**Is modifying the Dockerfile the only necessary step to add `patchwork` to the container?**

<details><summary>Answer</summary>
No, just because a Dockerfile has the package listed doesn't mean that our running container has that package.
</details>
::::
:::

### Planning containerization next steps

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**How does this package become usable for us?**:

Given what you know about the containerization process, what steps need to be done so that the `patchwork` package is available to use with the analysis scripts?

<details><summary>Answer</summary>

1. Build the modified Dockerfile to be an image (will need to do this using a command line interface)
2. Run the image to create a container (CLI or Docker Desktop), mounting the directory or volume that has the analysis scripts within the run command.
3. That running container now has `patchwork` available for use.

You could push the image after building it (in order to share it), but don't need to do that in order to use the package within a running container. Definitely share it eventually so others in your team have access to all of the same software you used!

</details>
::::
:::

We will perform these next steps of the containerization process in the Data Visualization section.

***
