# **Data Analysis**
***

If you have been following along but stopped, the scripts we are using are set up to both save and import the data to/from our {{< glossary "Mount" display="mounted" >}} {{< glossary "Volume" display="volume" >}} (the `ocs-bio-containers-scaffold-main` directory). Therefore in between sections, you should be able to pause, stop, or even delete a {{< glossary "Container" display="container" >}} and either unpause, re-start, re-create the container (Step 3b from the Data Import -- Container Infrastructure section), or create a new container. In this Data Analysis section, you will continue to use the new container created within the Data Visualization section. 

***

<details> <summary> If you skipped the Data Wrangling section or the Data Visualization section click here. </summary>

If you only skipped the Data Wrangling section and went through the Data Visualization section, you should be fine to proceed with the Data Analysis section.

If you only skipped the Data Visualization section but went through the Data Wrangling section, you will need to create the new container as was done in the Data Visualization -- Container Infrastructure steps.

If you skipped the data wrangling and the data visualization section, then you will need to fetch both the modified Dockerfile and the wrangled data, storing them in the `ocs-bio-containers-scaffold-main` directory from Step 3a in the Data Import -- Container Infrastructure section and you will need to create the new container as was done in the Data Visualization -- Container Infrastructure steps.

To fetch the modified Dockerfile and wrangled data, follow these instructions:

:::{.panel-tabset}

### Download from github.com

* Modified Dockerfile
    * Use this link to <a href="https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers/main/resources/dockerfiles/activity/Dockerfile" download> download the modified Dockerfile</a>.
    * Create a `dockerfiles` folder in the top directory of your `ocs-bio-containers-scaffold-main` folder (e.g., the one with `analysis_scripts`, `continued_learning_scripts`, the `README.md`)
    * Move the downloaded Dockerfile into the new `dockerfiles` folder. 
* Wrangled data
    * Use this link to [download the wrangled data as an `.rda` file](https://github.com/opencasestudies/ocs-bio-containers/raw/refs/heads/main/data/wrangled/wrangled_data.rda)
    * Create a `wrangled` folder within the `data` folder of `ocs-bio-containers-scaffold-main`.
    * Move the downloaded `.rda` file into the new `wrangled` folder.

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3fb50cd5a7c_0_246#slide=id.g3fb50cd5a7c_0_246")
```

### Command Line

* Ensure that you are in the `ocs-bio-containers-scaffold-main` top directory. 
    * If you type `pwd` (for print working directory), the `ocs-bio-containers-scaffold-main` should the final part of the printed path.
* Enter `mkdir -p dockerfiles data/wrangled`. 
    * This will create directories named `dockerfiles` in the top and `wrangled` under the `data` directory.
* Enter `cd dockerfiles`
* Copy and paste the following:

```{.bash}
wget "https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers/refs/heads/main/resources/dockerfiles/activity/Dockerfile" .
```

* Enter `cd ../data/wrangled`
* Copy and paste the following: 

```{.bash}
wget "https://github.com/opencasestudies/ocs-bio-containers/raw/refs/heads/main/data/wrangled/wrangled_data.rda" .
```

* Enter `cd ../../`

:::

</details>
***

## Container Application

### Correlation 

The correlation between two variables is a statistical measure of their relationship -- how they occur or vary together. Correlation is often used to describe relationships between variables but does not provide any evidence towards a cause and effect relationship between the two. Correlations are described with a [**correlation coefficient**](https://www.khanacademy.org/math/statistics-probability/describing-relationships-quantitative-data/scatterplots-and-correlation/a/correlation-coefficient-review?referrer=share_link) which ranges from -1 to 1, describing both the strength of the relationship and the direction of the relationship. Correlation coefficients often accompany scatter plots (like the ones produced in the Data Visualization section of this case study).

#### Strength of the relationship

* Correlation coefficient values equal to 0 mean that two variables don't seem to exhibit a relationship at all.
* Correlation coefficient values close to 0 mean that the two variables exhibit a weak relationship.
* Values close to -1 or 1 mean that the two variables exhibit a strong relationship. The closer to -1 or 1, the stronger the relationship.
* Values equal to -1 or 1 mean that the two variables exhibit perfect correlation.

#### Direction of the relationship

* Negative correlation coefficients (< 0) mean that two variables are inversely related, moving in opposite directions. As one variable increases, the other decreases.
* Positive correlation coefficients (> 0) mean that two variables are directly related, moving in the same direction. As one variable increases, the other also increases.

::: {.column-margin}
Make a figure like this one from a previous case study: https://www.opencasestudies.org/ocs-bp-co2-emissions/#Correlation_coefficient
:::

#### Spearman Correlation

Often the Pearson method is used to find the correlation between two variables. However this is a parametric method that assumes that the variables both have Normal (or Gaussian distributions), an important assumption if the analysis utilizes a hypothesis test together with finding the correlation coefficient. Another method is the Spearman method which is non-parametric and does not make this normality (or any underlying distributional) assumption about the data. Assumption-wise, the Spearman correlation does assume that each pair of values are paired but independent from other pairs. The Spearman correlation is found by ordering values from least to greatest (or the ordinal level of the variables) and using their ranks rather than numerical values. In this way, the Spearman correlation is reporting on the strength and direction of a monotonic relationship rather than a linear one and it is not sensitive to outliers.

:::{.column-margin}
Include a visual comparing linear (constant rate of change) vs monotonic relationships
:::

Because we don't know the underlying distribution of our variables (we haven't looked at that) and from the scatter plot there are a couple of points that separate from the others, Spearman correlation seems like a good choice to assess the relationship between the variables in our scatter plots. We are assuming that samples in each dataset are unique to that dataset (e.g., a research participant is only part of the `KIM_ADENOMAS` and none of the other studies). This seems like a fairly likely assumption. If our datasets were studying a rare condition or disease this could be an unreasonable assumption. Given the data that we have, we may have to consult with the rest of the team to confirm the veracity of this assumption. However, since it's secondary data, we may not be able to confirm this across studies.

:::{.emphasis-block}
For a discussion on the limitations of correlation, consider this summary from a previous case study: [https://www.opencasestudies.org/ocs-bp-co2-emissions/#Limitations_of_Correlation](https://www.opencasestudies.org/ocs-bp-co2-emissions/#Limitations_of_Correlation)
:::

#### What correlation relationships do we want to measure or quantify?

Our scatter plots are directly comparing the number of paired samples to the number of HMDB annotated compounds or KEGG annotated compounds in the dataset. So that's two correlations:

1. Number of paired samples vs the Number of HMDB annotated compounds
2. Number of paired samples vs the Number of KEGG annotated compounds

We also observed earlier that the datasets are mostly located in the same region (e.g., top right, bottom left) when comparing between the two scatter plots, so we could also compare the correlation between the number of annotated compounds for each database:

3. Number of HMDB annotated compounds vs Number of KEGG annotated compounds

We can consider all of the datasets together, or only look at subsets of the datasets. Previously, when looking at the scatter plots we observed that non-longitudinal datasets seemed to separate from longitudinal studies. We also have some datasets whose research participants are children or infants and others whose research participants are adults. Therefore, we could filter our datasets a few ways:

A. Find the correlation between the variables considering all datasets in the curated data resource.<br>
B. Find the correlation between the variables considering only non-longitudinal studies from the curated data resource.<br>
C. Find the correlation between the variables considering only datasets with adult participants from the curated data resource.<br>

:::{.emphasis_block}

<details><summary>Which studies are non-longitudinal?</summary>

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}

**How can we find non-longitudinal studies in our wrangled dataset and how many are there?**

<details><summary>Hint</summary>
What are the possible values in the `longitudinal` column?
</details>

<details><summary>Answer</summary>

We can find the datasets or studies whose value in the `longitudinal` column is "No". There are 8 studies. 

```{r, echo = FALSE, fig.alt = "There are 8 datsets where the longitudinal column records No to represent that it is a non-longitudinal study."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3fb8f9fbd14_0_0#slide=id.g3fb8f9fbd14_0_0")
```

In R, we would use:

```{r, eval = FALSE}
table1_extract %>%
    filter(longitudinal == "No")
```

</details>
::::
:::

</details>

<details><summary>Which datasets have adult participants?</summary> 

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}

**How can we find datasets with Adult participants in our wrangled dataset? And how many are there?**

<details><summary>Hint</summary>
What are the possible values in the `cohort_description` column?
</details>

<details><summary>Answer</summary>

The `cohort_description` column is much more complex than the `longitudinal` column. However, certain cohorts are described as being "Infants...", "Children....", "...from children...", "Preterm infants....". We could exclude those cohorts. Then there are 10 datasets.

```{r, echo = FALSE, fig.alt = "There are 10 datasets where the cohort description does not include the mention of children or infants, representing adult cohorts."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3fb8f9fbd14_0_11#slide=id.g3fb8f9fbd14_0_11")
```

In R, we would use:

```{r, eval = FALSE}
table1_extract %>%
    filter(!str_detect(tolower(cohort_description),
                      "infants|children"))
```

We will explain this code later.

</details>
::::
:::

</details>

:::

We will therefore find 9 correlation coefficients: 

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}

**What will the 9 correlation coefficients represent?**

<details><summary>Hint</summary>

We are looking at 3 different comparisons for the variables in the datasets (1-3 above) and 3 different sets within the datasets (A-C above).

</details>

<details><summary>Answer</summary>

* A1: Correlation between number of paired samples and number of HMDB annotated compounds, considering all datasets within the curated data resource.
* A2: Correlation between number of paired samples and number of KEGG annotated compounds, considering all datasets within the curated data resource.
* A3: Correlation between number of HMDB annotated compounds and number of KEGG annotated compounds, considering all datasets within the curated data resource.
* B1: Correlation between number of paired samples and number of HMDB annotated compounds, considering only non-longitudinal studies within the curated data resource.
* B2: Correlation between number of paired samples and number of KEGG annotated compounds, considering only non-longitudinal studies within the curated data resource.
* B3: Correlation between number of HMDB annotated compounds and number of KEGG annotated compounds, considering only non-longitudinal studies within the curated data resource.
* C1: Correlation between number of paired samples and number of HMDB annotated compounds, considering only datasets with Adult participants within the curated data resource.
* C2: Correlation between number of paired samples and number of KEGG annotated compounds, considering only datasets with Adult participants within the curated data resource.
* C3: Correlation between number of HMDB annotated compounds and number of KEGG annotated compounds, considering only datasets with Adult participants within the curated data resource.

</details>

::::
:::

#### Correlation Coefficient Matrices

A correlation coefficient matrix is a square table that displays the pairwise correlations between multiple variables in a dataset. Because we are comparing three variables from our dataset (for each subset of our dataset that we consider (e.g., all data, non-longitudinal, adult)), each set of our results will be in a correlation coefficient matrix.

```{r, echo = FALSE, fig.alt = "Correlation coefficient matrices are square & symmetric with the diagonal being irrelevant ones."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3fb8f9fbd14_0_22#slide=id.g3fb8f9fbd14_0_22")
```

Correlation coefficient matrices are symmetric meaning that they provide redundant information: The correlation between Variable A and Variable B is the same as the correlation between Variable B and Variable A (Correlation is commutative, like addition -- changing the order of the inputs does not change the final result). The diagonal of the matrix (dashed diagonal line from the top left to the bottom right in the figure above) is always 1 because every variable is always perfectly positively correlated with itself.

Because these values are stored in a matrix, if we want to extract a specific correlation coefficient, we will have to provide two names or indices to the matrix -- the row and column. It doesn't matter what order we provide these in since the matrix is symmetric/redundant.

### Running the correlation analysis

We will use the following script to find the correlations that we described previously.

```{.R filename="04_analysis_metric.R"}
{{< include resources/analysis_scripts/04_analysis_metric.R >}}
```

:::{.emphasis_block}
<details><summary>What is this script doing? </summary>

Lines 1-4 of the script are setup lines including the script's shebang in line 1 and loading necessary packages in lines 3-4. The `here` package will be used to build input the file path on line 6. The `tidyverse` is used to pipe the data from function to function (with `%>%`) and for functions like `filter` from `dplyr`, and `select` from `dplyr` that we've used before, as well as `str_detect` from `stringr`.

Line 6 of the script loads the `table1_extract` variable that stores the wrangled Table 1 data that we extracted and wrangled previously throughout the Data Import and Data Wrangling sections.

Lines 8-19 provide an `roxygen` block documenting a function `find_cor_matrix()` that is defined in lines 20 - 51.

* Lines 8-10 provide descriptions of the overall function.
* Lines 12-15 provide descriptions of the two parameters that are used to control the functionality or behavior of this function (`input_df` -- the input dataset and `filter_description` to control how that dataset should be filtered).
* Line 19 describes what the function returns to the user.
* Line 20 names the function and the paramters that control its behavior, including specifying that the `filter_description` parameter has a default value "All Data".
* Line 23 checks the `filter_description` parameter value using the [base R function `stopifnot()`](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/stopifnot). This means that only those 3 values are allowed values for the function. Any other values and the function wouldn't know how to proceed correctly or what the user wants, so the function stops.
* Lines 26-36 use the actual value of the `filter_description` parameter to filter the input dataset (the A, B, C parts of the comparisons we enumerated earlier). 
    * If the `filter_description` is `"Non-longitudinal"`, then we use the `dplyr` `filter()` function to keep only rows from the input dataframe where the `longitudinal` column are equal to  (`==`) "No" (lines 26-29).
    * Instead if `filter_description` is `"Adult"`, then we use the `dplyr` `filter()` function to keep a different set of rows (lines 29-33):
        * First, using the [base R `tolower()` function](https://www.rdocumentation.org/packages/h2o/versions/2.8.4.4/topics/tolower), we change every letter in the `cohort_description` column of the input dataframe to be lowercase. 
        * Then the completely lowercase column is given to the [`stringr` `str_detect()` function](https://stringr.tidyverse.org/reference/str_detect.html) to search for rows that have the words "infants" or (`|`) "children" somewhere in that column/row's intersection or value.
        * Instead of keeping the rows that have the "infants" or "children" strings, we're going to keep all of the other rows, this is accomplished by using the exclamation point (`!`) in front of the call to `str_detect()`
    * Finally, if `filter_description` didn't equal either "Non-longitudinal" or "Adult", then it must equal "All Data" because we already checked in line 23 that it was one of those 3. So in that case we just keep the whole dataset (lines 33-36).
* Lines 39-43 now run the creation of the correlation coefficient matrix (`cor_matrix`):
    * Lines 39-42 uses the filtered dataset, passing it to the `dplyr` `select` function to keep only the `num_paired_samples`, `kegg_annotated_compounds`, and `hmdb_annotated_compounds` column that we discussed earlier (the 1, 2, 3 parts of the comparisons we enumerated earlier).
    * This is then passed to the [`stats` `cor()` function](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/cor) (which is available through base R) on line 43.
        * The `.` specifies that the piped data frame is the input. Depending on the parameters given to the function when it's called this will have the exact number of datasets included (rows) we want and only 3 columns.
        * The `method` argument is used to specify that the Spearman correlation should be found.
* Lines 46-47 output information about the correlations such as which filter was used (All Data, Non-longitudinal studies, Adult cohorts) and the number of datasets per variable (e.g., the number of rows of the `to_assess` data frame following filtering).
* Lines 48-50 output correlation coefficients for each of the 1, 2, 3 comparisons.
    * First, the correlation between the number of KEGG annotated compounds and the number of HMDB annotated compounds.
    * Second, the correlation between the number of paired samples and the number of KEGG annotated compounds.
    * Finally, the correlation between the number of HMDB annotated compounds and the number of paired samples.

Lines 54, 57, and 60 call the `find_cor_matrix()` function, each for a different subset of the datasets from the curated data resource.

* Line 54 uses all datasets
* Line 57 uses non-longitudinal studies
* Line 60 uses datasets with adult cohorts.


Notice that this script is outputing messages like previous scripts have done, but it is not saving any files. **Therefore, we will have to be intentional in how we capture the output from it when we run it.**

</details>
:::

Knowing that this script is in the `analysis_scripts` directory and is an `.R` like all of the other scripts that we have run throughout this case study, enter a command to run this script in your running container.

```suggestion

::: {.emphasis_block}
<details><summary>Reminder of the correct command:</summary>

Making sure that you're in the appropriate directory use either: 

```{.bash}
Rscript 04_analysis_metric.R
```

or

```{.bash}
Rscript analysis_scripts/04_analysis_metric.R
```

</details>
:::
```

### Reviewing the output

Notice that (most likely) the output of the script is printed to the Docker Desktop or CLI console, but isn't stored in a file. You could potentially copy/paste them into a document, or take a screenshot, but it would be better, if we captured the output in a file within our mounted volume when we run the script.

### Running the analysis again

This time, use the output redirection operator (`>`) when you run the script to redirect the output from the script into a file. Consider [this article for more information about output redirection](https://www.geeksforgeeks.org/linux-unix/input-output-redirection-in-linux/).

:::{.panel-tabset}

#### Docker Desktop

* On the "Exec" tab of the running container, ensure that you are in the `analysis_scripts` directory of the mounted volume.
* Enter the following command:

```{.bash}
mkdir -p ../results/logs
```

This creates a directory within the `results` directory called "logs" and will create the `results` directory (on the same level as `analysis_scripts`) if it doesn't already exist.

* Enter the following command:

```{.bash}
Rscript 04_analysis_metric.R > ../results/logs/correlation_result.txt
```

This runs the script, redirecting its standard output (our messages and correlation coefficients) to a text file named `correlation_result.txt` within the directory you just created.

* Enter the following command to check the contents of that file:

```{.bash}
cat ../results/logs/correlation_result.txt
```


#### Docker through a CLI

* Ensure that you are in the root directory of your mounted volume.
* Enter the following command to create a new subdirectory for output in the `results` directory:

```{.bash}
mkdir -p results/logs
```

* Enter the following command to run the script and redirect the standard output from the script to a file in that new directory:

```{.bash}
Rscript analysis_scripts/04_analysis_metric.R > results/logs/correlation_result.txt
```

* Enter the following command to view the contents of that file:

```{.bash}
cat results/logs/correlation_result.txt
```

:::

::: {.emphasis_block}

```{r}
#| echo: false
#| output: asis

# Replace with your raw GitHub URL
url <- "https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers-analysis/refs/heads/main/analysis_results/logs/analysis_logs.txt"

# Fetch lines and print verbatim as Markdown with two spaces in between lines to enforce lines breaks
file_lines <- readLines(url, warn = FALSE)
formatted_lines <- paste0(file_lines, "  ")
cat(formatted_lines, sep = "\n")
```

:::

:::{.think_question}
::::{.think_question_header}
Question Opportunity
::::
::::{.inner_block}
**What do you observe about the correlation coefficients?**

<details><summary>Some observations</summary>

* The correlation between the number of annotated compounds for the HMDB and KEGG reference databases is the highest. It isn't always 1 (perfect positive correlation), but it is a very strong monotonic correlation.
* The correlation between the number of paired samples and the number of annotated compounds from a reference database is weaker. It's also not always equal if looking at number of paired samples compared to KEGG vs the number of paired samples compared to HMDB. 
    * This could be due in part to the dataset that we called out earlier `iHMP_IBDMDB_2019`. It is a longitudinal dataset so it would not be in the non-longitudinal dataset correlation analysis where now the correlation coefficients match (0.45). From this, we might hypothesize that the ordinal rankings of the non-longitudinal datasets are the same whether using the number of KEGG annotated compounds or the number of HMDB annotated compounds.
</details>
::::
:::


### Team Collaboration

An analysis changes over time as you develop it. Even with numbering our scripts and using a container, team members may have a difficult time reproducing our analysis unless they know exactly what was run in what order.

:::{.reproducibility_box}
::::{.reproducibility_box_header}
Reproducibility considerations
::::
::::{.inner_block}
Having a "run all" script that specifies what scripts were used and in what order for the analysis is a best practice for reproducibility [@Wilson_etal_2017].
::::
:::

For this analysis, a run all script might look like the following:

```{.bash filename="run_analysis.sh"}
{{< include resources/analysis_scripts/run_analysis.sh >}}
```

:::{.emphasis_block}
<details><summary>What's this script doing?</summary>

This is not an R script like the other scripts have been throughout this case study. Instead this is a bash script. Every `cd`, `ls`, `pwd`, `docker build`, `Rscript [script name]`, `mkdir` command that you've used through out this case study is a "bash shell" command. Therefore this bash script is recording those commands in the correct order that they need to happen.

Line 1 of the script is the script's shebang. Since this is a bash script, it has a different shebang than an R script would have.

Lines 3, 9, 13, 19, and 23 all are messages describing where the run all process is / what it will be doing next. This is similar to the `cat` messages we had in the R scripts.

Line 5 makes the `results/logs` directory like we did previously. All other directories are made by the scripts if they don't exist.

Line 7 calls the first script that we used in the Data Import section and redirects the output messages to a log file.

Line 11 calls the second script that we used in the Data Wrangling section and redirects the output messages to another log file.

Line 15 calls the first data visualization script and redirects the output messages to another new log file.

Line 17 calls the second data visualization script and redirects its output messages to the same log file used for the first script. This time the `>>` redirection operator is used. This appends to a pre-existing file rather than overwriting it.

Line 21 calls the final R script with the correlation analysis and redirects its output messages to yet another new log file.

Finally, line 25 is an inline command. Rather than saving everything that's in the quotes (`"`) inside of a script, we include it here and tell the Bash script to execute the command as if it was in an R script. This also allows us to mix bash commands with R code. 

* The first portion that's inside the `$( ... )` is the bash part
    * [`cat`](https://www.geeksforgeeks.org/linux-unix/cat-command-in-linux-with-examples/) (which we used earlier to look at the contents of a file) all of (`*`) the R scripts in this directory
    * send the output of that (`|` -- bash's version of a pipe like the `%>%` R had) to ...
    * [`grep`](https://www.geeksforgeeks.org/linux-unix/grep-command-in-unixlinux/) which will pull out only lines that have "library" in it
    * send the output of that (`|`) to ...
    * [`sort -u`](https://www.geeksforgeeks.org/linux-unix/sort-command-linuxunix-examples/) which will pull out only the unique (`-u`) lines. (It first sorts it for algorithmic efficiency.)
    * send the output of that (`|`) to ...
    * a bash function to take a stream of newline separated strings and format them into a single line separated by a semicolon and space. This allows us to pack a multi-step sequence into a single inline command. 
* We finally combine this with a [`sessionInfo()`](https://sessioninfo.r-lib.org/) call (which is an R function)
* And redirect the output of the whole thing into a log file
* The purpose of that whole line was to output package versions and other session info for the R packages that we used in the analysis!

</details>
:::

Point out that the validator can get the same results as the learner did by running the `run_analysis.sh` script.

::: {.panel-tabset}

### Docker Desktop

Ask learners to try to run the analysis

### Docker through a CLI

Ask learners to try to run the analysis

:::



## Container Infrastructure

In this section we can discuss how someone else (maybe the person validating the work for resubmission) can use the container to run the analysis and get the exact same results and plot that you made because of the container. Specifically, what should you share with someone so they can rerun the analysis?

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f0102bcf57_7_5#slide=id.g3f0102bcf57_7_5")
```



Sharing the container image vs sharing the Dockerfile



Add a reproducibility box about pushing the new / adjusted scripts and plots as well as the documentation to the scaffold repo that they "used as a template/cloned" if they did the earlier extra reproducibility step.



***
