# **Data Import**
***

## Container

Activity: access the container the collaborator shared with you (because this is how you'll be able to get the data extracted from the PDF import).

Side quest: find a container on dockerhub (and discuss reputable sources)

Note that although we can see the container -- this doesn't give us access to the data analysis data yet. For that we must build the image and run the container.

## Data analysis data

Using the tabulapdf package, we will extract a raw, unwrangled version of Table 1.

Walk through how to run the container in order to run the analysis and extract the data.

Side quest: what if you wanted to spin up an IDE with your container?

Activity -- Using the container from the collaborator to get the Table 1 data.

Because we want to avoid mounting data and scripts in this case study (and leave that as a side quest/suggested hw), we'll use the following instructions for someone to interact with the data analysis data within a running container:

For Docker Desktop:

* grab the container from docker hub
* run the container
* navigate to the exec tab
* `touch` an `analysis.R` script in `opt/case_study_analysis`, a folder the Dockerfile specifies creation of
* navigate to the files tab
* right click on the `analysis.R` script and click edit
* copy paste the following code into the script
* click save
* navigate to the exec tab
* use `Rscript analysis.R` -- adjusting based on script location and learner's file location.
* notice that it outputs "Dimensions: 25 x 7"
* notice that there is an rda file in the `data/imported` folder

```{r, eval = FALSE}
#!/usr/bin/env Rscript

raw_table <- tabulapdf::extract_tables("https://www.nature.com/articles/s41522-022-00345-5.pdf",
                                       pages = 2,
                                       method = "stream",
                                       col_names = FALSE
                                      )[[1]]

message("Dimensions: ", paste(dim(raw_table), collapse = " x "))
```

```{r, eval = FALSE}
save(raw_table, file = here::here("data", "imported", "raw_table1.rda"))
```

***
