# **Data Wrangling**
***

## Container Application

If you have been following along but stopped, the scripts we are using are set up to both save and import the data to/from our mounted volume (the `ocs-bio-containers-scaffold-main` directory). Therefore in between sections, you should be able to pause, stop, or even delete a container and either unpause, re-start, or re-create the container (Step 3b from the Data Import section). 

::: {.column-margin}
Add instructions on pausing, stopping, and deleting containers.
:::

***

<details> <summary> If you skipped the data import section click here. </summary>

At the minimum, you will need to repeat the Container Infrastructure steps from the Data Import section. 

You can either perform Step 1 from the Data Import -- Container Application section, or in your now running container, you can use `wget` to fetch the data. 

If using `wget` to fetch the data:

:::{.panel-tabset}

### Docker Desktop

* Navigate to the "Exec" tab (along the top of the running container).
* Type `cd /home` and press the return / enter key.
* Type `mkdir -p data/imported` and press the return / enter key.
* Type `cd data/imported` and press the return / enter key.
* Copy and paste the following: 

```{.bash}
wget "https://github.com/opencasestudies/ocs-bio-containers/raw/refs/heads/main/data/imported/raw_table1.rda" .
```

* Type `cd /home` and press the return / enter key.

### Docker through a CLI 

* Enter `mkdir -p data/imported`
* Enter `cd data/imported`
* Copy and paste the following: 

```{.bash}
wget "https://github.com/opencasestudies/ocs-bio-containers/raw/refs/heads/main/data/imported/raw_table1.rda" .
```

* Enter `cd ../../`

:::

</details>
***

### Wrangling the data

Using the `tidyverse`, we will wrangle the extracted version of Table 1. This script is provided for you as part of scaffold code in the `ocs-bio-containers-scaffold-main` directory that was mounted as a volume when running the image to create the container.

```{.R filename="02_wrangle.R"}
{{< include resources/analysis_scripts/02_wrangle.R >}}
```

:::{.emphasis_block}

<details><summary>What is this script doing?</summary>

Lines 1-5 of the script are setup lines including the script's shebang in line 1 and loading necessary packages in lines 3-5. `here` will be used to build input and output file paths on lines 7, 44, 50, and 53. `tidyverse` is used to pipe the data from function to function and for functions like `rename`, `mutate`, `filter`, and `select` throughout the script. `readr` will be used to write an output `csv` file on line 52.

Line 7 loads the `raw_table` variable that stores the Raw Table 1 data that we extracted and saved using `tabulapdf` in the Data Import Section. 

Lines 10-36 are the actual wrangling steps of the process.

* Lines 11-16 rename the data frame columns so they are descriptive, reflecting the data they contain.
* Lines 17-22, 32 extract the numbers (which represent the number of paired samples) that were appended to the ends of the cohort descriptions from the `cohort_description_num_paired_samples` column, storing them in their own column `num_paired_samples`.
* Lines 17, 23-26, 32 remove the numbers that we just extracted from the `cohort_description_num_paired_samples` column, creating an intermediate text column (named `cohort_description_text`) that only contains the cohort descriptions (though they are still incomplete for a given row because they are multi-line).
* Lines 17, 27-31, 32 sets up a `case_when` that evaluates whether the next value/row in the dataset name column is an NA value. 
    * This utilizes the `lead` function to look for the next value. 
    * In the case of the last row, we're using the `default` argument to pad the data with a `FALSE`. 
    * If the `lead` function returns that the next value is NA (`TRUE`), it will paste the current cohort text with the next row's text (the text that we stored in the intermediate `cohort_description_text` column in the previous step). 
    *. Otherwise (e.g., if the `lead` function returns that the next value is not NA (`FALSE`) or is the last value/row), it will return the current cohort text only.
    * At the end of this, the `cohort_description` column contains complete, single line cohort descriptions. 
* Line 33 is a filter to remove the rows where the dataset name is NA because we've already appended the important information from those rows to the previous rows and we no longer need them. Specifically, the `filter` function will only keep rows where the dataset name is not NA.
* Lines 34 - 36 remove 
    * the `X4` column which was all `NAs`
    * `cohort_description_num_paired_samples`: the original third column (that had the cohort description plus number of paired samples), 
    * `cohort_description_text`: the intermediate column which we made that has incomplete cohort descriptions

The rest of the script is used to output information about the wrangled data or save the wrangled data itself:

* Line 39 ouputs the dimensions of the wrangled table.
* Line 40 outputs the number of NA values that are in the wrangled table.
* Line 41 extracts a specific dataset name from row 4 of the table.
* Line 44 specifies an output directory.
* Line 46 checks to see if that output directory exists, and if not creates it in line 47.
* Lines 50-51 save the wrangled data as an `.rda` file.
* Lines 52-54 save the wrangled data as a `.csv` file.

</details>

:::

::: {.panel-tabset}

To run this script in the running container in order to wrangle the data, follow these steps:


### Docker Desktop

* Verify that you are in the "Exec" tab of running container. 
* Verify which directory you are in by typing `pwd` and pressing the enter / return key.
* Navigate to the `/home/analysis_scripts` directory using `cd` (e.g., `cd analysis_scripts` if in `home`).
* Enter the following command and press the return / enter key.

```{.bash}
Rscript 02_wrangle.R
```

* Notice the output messages from the script.

### Docker through a CLI

Verify that you are in the top directory (`ocs-bio-containers-scaffold-main`) and Enter the following command in your CLI.

```{.bash}
Rscript analysis_scripts/02_wrangle.R
```

Notice the output messages from the script.

:::

### Verifying the wrangled data

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**What does the output from the messages in the script tell us about our wrangled data?**:

Certain lines have certain information about the data -- what are they reporting and what do we expect them to say?

1. The dimensions of the data frame
2. The number of NAs within the data frame
3. The dataset name for the fourth row of data (spot checking a value)

<details><summary>Answer</summary>

If we were to inspect the `table1_extract` Data Frame, we would observe that the data now appears to match our expectations. 

```{r}
#| echo: false
#| warning: false
#| message: false

url <- "https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers-analysis/refs/heads/main/wrangle_data/wrangled/wrangled_data.csv"

table1_extract <- read_csv(url)
table1_extract %>% print(n=14)

```

There appear to be 14 rows and 7 columns, though those columns don't follow the original order of Table 1 in @Muller_Algavi_Borenstein_2022. Since the columns have descriptive names though, it will be fine that they are not in the same order. While the columns may not be in the same order as the source data, the rows should be in the same order as the source data. Spot checking a specific value will help us to verify this.

```{r wrangled_data, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = 'The wrangled data appears to match our expectations.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f5db0d7ede_0_0#slide=id.g3f5db0d7ede_0_0")
```

The log messages report the following:

```{r}
#| echo: false
#| output: asis

# Replace with your raw GitHub URL
url <- "https://raw.githubusercontent.com/opencasestudies/ocs-bio-containers-analysis/refs/heads/main/wrangle_results/logs/wrangling_logs.txt"

# Fetch lines and print verbatim as Markdown with two spaces in between lines to enforce lines breaks
file_lines <- readLines(url, warn = FALSE)
formatted_lines <- paste0(file_lines, "  ")
cat(formatted_lines, sep = "\n")
```

The dimensions properly report a decrease in the number of rows compared to the raw extracted data (14 instead of 25), matching our expectation for both the number of rows and the number of columns given the source data. The dataset name in the fourth row is `HE_INFANTS_MFGM_2019`, again matching our expectations when comparing to the source data.

Log message purpose | Expected | Observed
:-------------------|:---------|:--------
New dimensions      | 14 x 7   | 14 x 7
New number of NAs   | 0        | 0
Specific value check| `HE_INFANTS_MFGM_2019` | `HE_INFANTS_MFGM_2019`

</details>
::::
:::

We can verify that the wrangled data was saved by following the same process that we did in the Data Import -- Container Application Step 2 section above. 

:::{.think_question}
::::{.think_question_header}
Advanced Question opportunity
::::
::::{.inner_block}
**A more generalizable and robust wrangling approach**:

This wrangling approach using the `lead` function to assess if there is multiline data that needs combined assumes that there are only ever 1 - 2 rows for each cohort description. What changes to this process could be utilized that would be more generalizable or robust with data that doesn't meet this assumption? Specifically, how could this process be changed to combine an unknown number of rows of data to the previous row when only one column has data?

<details><summary>Answer</summary>
One possible approach would:

1. Use the `cumsum()` function to assign a group ID to the rows.
2. Use the `group_by()` function to group the data by that group ID.
3. With the `summarize()` function, `paste()` the cohort descriptions within each group together.
4. Either bring along the non-NA data for the other columns within the `summarize()` function, or join the datasets together using the group ID and drop unnecessary rows.

</details>
::::
:::

### Planning data visualizations

Now that we've wrangled the extracted table data into a Data Frame, we want to plan out the visualizations we'll make using this Data Frame!

Collab O. Rator's email provided the following guidance regarding the data visualizations:

>Specifically, I’d like to see scatter plots comparing the data from the final four columns for each dataset (No. samples w/ paired data, Longitudinal, No. HMDB-annotated compounds, and No. KEGG-annotated compounds).

Collab O. Rator's instructions have asked us to visualize 4 variables or dimensions of the data with scatter plots. The axes (x- and y-) will account for two of those variables, and we can adjust the points (e.g., color or shape) to represent a third variable -- likely the `longitudinal` value as this is categorical ("Yes" vs "No"). To visualize the fourth variable, we would need to add a third axis (z-) or some other means of encoding the data, but this adds to the complexity of the plot. Instead, we could make two separate panels.

Plot element | Data column
-------------|------------
x-axis       | `num_paired_samples`
y-axis       | `hmdb_annotated_compounds` or `kegg_annotated_compounds`
point color or shape | `longitudinal`
labels               | `dataset_name`

But if we make two separate scatter plots, they will be two separate plots/files that won't be displayed next to each other unless we use design software like Illustrator to combine them later. Certain R packages allow you to combine plots and display them with each other. So let's "wrangle" the Dockerfile to include a package that will let us do that: `patchwork`.

:::{.emphasis_block}

<details><summary>Why can't we just run the data visualization scripts in our current running container?</summary>

Our current running container has several packages available to us, but `patchwork` is not one of them. If you want to verify this, type `Rscript -e "library(patchwork)"` in your running container (specifically in the "Exec" tab for Docker Desktop). You will see an error message.

>Error in library(patchwork) : there is no package called ‘patchwork’
Execution halted

</details>

:::

## Container Infrastructure

As you can see from the visualization planning above, these two plots would be completely separate -- and we'd like them to be visible next to each other. A useful R package for this purpose is `patchwork`. However, our data visualization Docker container doesn't include that package.

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}

**How can we add a package to a container?**

<details><summary>Answer</summary>

There are multiple ways to add a package to a container.

If the container is running, we could install the package, and it will be usable/available, while we are using the running container (or if we pause and later restart that container). However, that package won't be available to someone else who is using the same image that we used to create the container. 

Another way to add a package to a container is to go back to the beginning definition of or recipe for that container (a Dockerfile), modify it, and then rebuild an image, and run the rebuilt image so that a new container has the additional package. 


Approach      | Pro  | Con
:-----------  | :---:|:---:
Install package in running container | Fast and we can continue our work quickly | Package not available to others using the same image we did
Modify the Dockerfile, then build, share, and run an image | All software we used is defined and available to others  | Longer containerization process before we can continue our work 


</details>

::::
:::

Wanting to make sure that we're practicing reproducibility best practices when working with containers, especially in a team environment, we're going to add a package to our container by modifying the Dockerfile, rebuilding the image, and running the container from the new image.

::: {.emphasis_block}

<details><summary>How would you install an R package from CRAN within a running container?</summary>

In the running container (specifically in the "Exec" tab for Docker Desktop), you could use `Rscript -e "install.packages('patchwork')"` to install the `patchwork` package -- replacing 'patchwork' with any other package availabe on CRAN. 

The [examples provided in this "Containers for Scientists" course from the ITN](https://hutchdatascience.org/Containers_for_Scientists/05-modifying-containers.html#templates-for-adding-packages) show different templates for how to specify the installation of different kinds of packages within a Dockerfile. If you remove the "RUN" from the example/template, the code will work to install the packages in a running container (as long as the container has Python, etc. already)

**It is best practice to modify the Dockerfile that specifies the image used to create the container in order to add package(s) to a container.**

</details>

:::

### Adding a package

Now we will add `patchwork` to the container -- or at least get started with the process of adding the `patchwork` package to the container.

Copy the [Dockerfile from the Data Exploration -- Container Infrastructure section](https://www.opencasestudies.org/ocs-bio-containers/#container-infrastructure-1) above. With your favorite text editor (e.g., RStudio, Visual Studio Code, Notepad++, etc.), modify the Dockerfile to also install the `patchwork` package.

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**What steps do you need to do to add `patchwork` as a package that will be installed and available within the eventual running container?**

<details><summary>Hint</summary>

Several CRAN packages are already being installed in this Dockerfile. You can just add `patchwork` to the list.

</details>

<details><summary>Answer</summary>

1. Add a space and backslash (` \`) to line 32 after the `gt` package.
2. Add `patchwork` to line 33, indented the same way that the rest of the packages in that section are.

</details>
::::

Save the modified Dockerfile within your `ocs-bio-containers-scaffold-main` directory from Step 3a in the Data Import -- Container Infrastructure section: 

* Create a directory called `dockerfiles` within the top layer of this directory (e.g., on the same level as `analysis_scripts`, `data`, `results`, etc.).
* Save the modified Dockerfile within this new `dockerfiles` directory, naming it `Dockerfile` exactly.

:::{.emphasis_block}
<details><summary>Your modified Dockerfile should look like...</summary>

```{.dockerfile filename="Dockerfile"}
{{< include resources/dockerfiles/activity/Dockerfile >}}
```

</summary>
:::

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**Is modifying the Dockerfile the only necessary step to add `patchwork` to the container?**

<details><summary>Answer</summary>
No, just because a Dockerfile has the package listed doesn't mean that our running container has that package.
</details>
::::
:::

### Planning containerization next steps

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**How does this package become usable for us?**:

Given what you know about the containerization process, what steps need to be done so that the `patchwork` package is available to use with the analysis scripts?

<details><summary>Answer</summary>

1. Build the modified Dockerfile to be an image (will need to do this using a command line interface)
2. Run the image to create a container (CLI or Docker Desktop), mounting the directory or volume that has the analysis scripts within the run command.
3. That running container now has `patchwork` available for use.

You could push the image after building it (in order to share it), but don't need to do that in order to use the package within a running container. Definitely share it eventually so others in your team have access to all of the same software you used!

</details>
::::
:::

We will perform these next steps of the containerization process in the Data Visualization section.

:::{.column-margin}
Add instructions about container clean up for the starting container that doesn't have patchwork.
:::

***
