<!--
Update the file paths as needed throughout
-->

<!--
At the end of this section, save incremental data so that users can pause and come back later
-->

# **Data Import**
***

## Container

Activity: access the container the collaborator shared with you (because this is how you'll be able to get the data extracted from the PDF import).

Side quest: find a container on dockerhub (and discuss reputable sources)

Note that although we can see the container -- this doesn't give us access to the data analysis data yet. For that we must build the image and run the container.

## Data analysis data

Using the tabulapdf package, we will extract a raw, unwrangled version of Table 1.

Walk through how to run the container in order to run the analysis and extract the data.

Side quest: what if you wanted to spin up an IDE with your container?

```{r, eval = FALSE}
raw_table <- tabulapdf::extract_tables("https://www.nature.com/articles/s41522-022-00345-5.pdf",
                                       pages = 2,
                                       method = "stream",
                                       col_names = FALSE
                                      )[[1]]
```

<!--
At the end of the data import section, save the imported data as .rda files
to allow users to stop and pick up where they left off:
-->

```{r, eval = FALSE}
save(raw_table, file = here::here("data", "imported", "raw_table1.rda"))
```

<!--
NOTE that depending on the size of the data for Bio-OCS, the data may not be in the case study project folder directly
and may need to be imported from an external source
-->

***
