<!--
Update the file paths as needed throughout
-->

<!--
Some case studies may wish to merge Data Exploration with this Wrangling subsection
-->

<!--
At the end of this section, save incremental data so that users can pause and come back later
-->

# **Data Wrangling**
***

If you have been following along but stopped, we could load our imported data like so:

```{r, eval = FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```
***

<details> <summary> If you skipped the data import section click here. </summary>

An RDA version (stands for R data) of the data can be found [here](https://github.com//opencasestudies/ocs-bp-co2-emissions/tree/master/data/imported) or slightly more directly [here](https://raw.githubusercontent.com/opencasestudies/ocs-bp-co2-emissions/master/data/imported/co2_data_imported.rda). Download this file and then place it in your current working directory within a subdirectory called "imported" within a directory called "data" to use the following code. We used an RStudio project and the [`here` package](https://github.com/jennybc/here_here) to navigate to the file more easily.

```{r, eval=FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```

</details>
***

```{r, eval = FALSE}

table1_extract <- raw_table %>%
    mutate(num_paired_samples = as.numeric(str_extract(X3, "\\d+$"))) %>%
    mutate(X3Text = str_remove(X3, "\\d+$")) %>%
    mutate(cohort_description = if_else(lead(is.na(X1), default = FALSE),
                                        paste(X3Text, lead(X3Text), sep = ""),
                                        X3Text)
           ) %>%
    filter(!is.na(X1)) %>%
    select(!c(X3, X4, X3Text)) %>%
    rename(c("dataset_name" = "X1",
             "ref" = "X2",
             "longitudinal" = "X5",
             "hmdb_annotated_compounds" = "X6",
             "kegg_annotated_compounds" = "X7"))

```

To allow users to skip import and wrangling we will save the data as an RDA file as well as a CSV file as this is often useful to send our data to collaborators. We will save this in a “wrangled” subdirectory of our “data” directory of our working directory.

```{r, eval = FALSE}
save(table1_extract, file = here::here("data", "wrangled", "wrangled_data.rda")) # <1>
readr::write_csv(table1_extract, file = here::here("data","wrangled",            # <2>
                                       "wrangled_data.csv"))         # <2>
```

1. saving the data as an RDA file within the wrangled data subdirectory
2. saving the data as a CSV file within the wrangled data subdirectory

<!--
Note that with this code annotation example, we've used a new line to space the code so that the annotation doesn't overlay the code
-->


***
