# **Data Wrangling**
***

## Container Application

If you have been following along but stopped, the scripts we are using are set up to load our imported data like so within our analysis script within the running container (without having to run the original `extract_tables` code). make sure your container is running (probably will need a tab structure for this): <- update to be more precise

```{r, eval = FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```
***

Even if you stop the container (as long as you don't delete it), you can restart it and the data will still be there as long as it was saved during wrangling

<details> <summary> If you skipped the data import section click here. </summary>

KI note: We'll need to add instructions to this about mounting or importing data if someone wants to pick up at this section. I don't think the container has `wget`, but I could set up a download that would have the file that they could then mount. Will do that and include instructions here for that.

```{r, eval=FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```

</details>
***

### Wrangling the data

::: {.panel-tabset}

We have a file that is part of the scaffold code that will wrangle the data in the ways we need.

```{.R filename="02_wrangle.R"}
{{< include resources/analysis_files/02_wrangle.R >}}
```


1. Rename the columns so they are descriptive
2. Extracting the numbers (which represent the number of paired samples) at the end of the cohort descriptions to put them in their own column
3. Making a intermediate column, specifically a version of the text column that has those numbers of paired samples removed
4. Setting up a `case_when` that evaluates whether the next value/row in the dataset name column is an NA value. This utilizes the `lead` function to look for the next value. In the case of the last row, we're using the `default` argument to pad the data with a `FALSE`
5. If the `lead` function returns that the next value is NA (`TRUE`), it will paste the current cohort text with the next row's text
6. Otherwise (e.g., if the `lead` function returns that the next value is not NA (`FALSE`) or is the last value/row), it will return the current cohort text only.
7. This filters to remove the rows where the dataset name is NA because we've already appended the important information from those rows to the previous rows
8. This removes the `X4` column which was all `NAs`, the original third column (that had the cohort description plus number of paired samples), and the intermediate column which we made that has incomplete cohort descriptions

Discuss what the above script is actually doing.

### Docker Desktop

how to run analysis in running container

### Docker through a CLI

how to run analysis

:::

### Verifying the wrangled data

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**What does the output from the messages in the script tell us about our wrangled data?**:

Certain lines have certain information about the data -- what are they reporting and what do we expect them to say?

1. The dimensions of the data frame
2. The number of NAs within the data frame
3. The dataset name for the fourth row of data (spot checking a value)

<details><summary>Answer</summary>

If we were to inspect the `table1_extract` Data Frame, we would observe that the data now appears to match our expectations. There are 14 rows and 7 columns, though those columns don't follow the original order of Table 1 in @Muller_Algavi_Borenstein_2022. Since the columns have descriptive names though, it will be fine that they are not in the same order. While the columns may not be in the same order as the source data, the rows should be in the same order as the source data.

```{r wrangled_data, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = 'The wrangled data appears to match our expectations.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f5db0d7ede_0_0#slide=id.g3f5db0d7ede_0_0")
```

The dimensions properly report a decrease in the number of rows compared to the raw extracted data (14 instead of 25), matching our expectation for both the number of rows and the number of columns given the source data. The dataset name in the fourth row is `HE_INFANTS_MFGM_2019`, again matching our expectations when comparing to the source data.

Log message purpose | Expected | Observed
:-------------------|:---------|:--------
New dimensions      | 14 x 7   | 14 x 7
New number of NAs   | 0        | 0
Specific value check| `HE_INFANTS_MFGM_2019` | `HE_INFANTS_MFGM_2019`

</details>
::::
:::

To allow learners to skip import and wrangling in later sections, we will save the data as an RDA file as well as a CSV file as this is often useful to send our data to collaborators. We will save this in a “wrangled” subdirectory of our “data” directory of our working directory.

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**A more generalizable and robust wrangling approach**:

This wrangling approach using the `lead` function to assess if there is multiline data that needs combined assumes that there are only ever 1 - 2 rows for each cohort description. What changes to this process could be utilized that would be more generalizable or robust with data that doesn't meet this assumption? Specifically, how could this process be changed to combine an unknown number of rows of data to the previous row when only one column has data?

<details><summary>Answer</summary>
One possible approach would:

1. Use the `cumsum()` function to assign a group ID to the rows.
2. Use the `group_by()` function to group the data by that group ID.
3. With the `summarize()` function, `paste()` the cohort descriptions within each group together.
4. Either bring along the non-NA data for the other columns within the `summarize()` function, or join the datasets together using the group ID and drop unnecessary rows.

</details>
::::
:::

### Planning data visualizations

Now that we've wrangled the extracted table data into a Data Frame, we want to plan out the visualizations we'll make using this Data Frame!

Collab O. Rator's email provided the following guidance regarding the data visualizations:

>Specifically, I’d like to see scatter plots comparing the data from the final four columns for each dataset (No. samples w/ paired data, Longitudinal, No. HMDB-annotated compounds, and No. KEGG-annotated compounds).

Collab O. Rator's instructions have asked us to visualize 4 variables or dimensions of the data with scatter plots. The axes (x- and y-) will account for two of those variables, and we can adjust the points (e.g., color or shape) to represent a third variable -- likely the `longitudinal` value as this is categorical ("Yes" vs "No"). To visualize the fourth variable, we would need to add a third axis (z-) or some other means of encoding the data, but this adds to the complexity of the plot. Instead, we could make two separate panels.

Plot element | Data column
-------------|------------
x-axis       | `num_paired_samples`
y-axis       | 'hmdb_annotated_compounds' or `kegg_annotated_compounds`
point color or shape | `longitudinal`
labels               | `dataset_name`

But if we make two separate scatter plots, they will be two separate plots/files that won't be displayed next to each other unless we use design software like Illustrator to combine them later. Certain R packages allow you to combine plots and display them with each other. So let's "wrangle" the Dockerfile to include a package that will let us do that: `patchwork`.

Could potentially have learners run the data visualization script that produces two separate data visualizations, but I think that's best kept for the next section?

## Container Infrastructure

As you can see from the visualization planning above, these two plots would be completely separate -- and we'd like them to be visible next to each other. A useful R package for this purpose is `patchwork`. However, our data visualization Docker image doesn't include that package.

### Adding a package

Activity: Add `patchwork` to the container

* edit the Dockerfile (in your favorite text editor)

### Planning containerization next steps

:::{.think_question}
::::{.think_question_header}
Question opportunity
::::
::::{.inner_block}
**How does this package become useable for us?**:

Rephrase this question, but basically ask the learners to consider info from the Introduction to Containers and containerization process and plan what will need to be done so that this package is available to use by the analysis scripts within a running container

<details><summary>Answer</summary>

build the Dockerfile to be an image (will need to do this on terminal) 

</details>
::::
:::


***
