# **Data Wrangling**
***

## Application to data

If you have been following along but stopped, we could load our imported data like so within our analysis script within the running container (without having to run the original `extract_tables` code):

```{r, eval = FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```
***

<details> <summary> If you skipped the data import section click here. </summary>

KI note: We'll need to add instructions to this about mounting or importing data if someone wants to pick up at this section.

```{r, eval=FALSE}
load(here::here("data", "imported", "raw_table1.rda"))
```

</details>
***

We'll need to add this code to our `analysis.R` file using the same steps as before:

* navigate to the files tab
* right click on the `preprocessing.R` script and click edit
* copy paste the following code into the script
* click save
* navigate to the exec tab
* use `preprocessing.R` -- adjusting based on script location and learner's file location.

table1_extract <- raw_table %>%
    rename(c("dataset_name" = "X1",                                              #<1>
             "ref" = "X2",                                                       #<1>
             "cohort_description_num_paired_samples" = "X3",                     #<1>
             "longitudinal" = "X5",                                              #<1>
             "hmdb_annotated_compounds" = "X6",                                  #<1>
             "kegg_annotated_compounds" = "X7")) %>%                             #<1>
    mutate(num_paired_samples = as.numeric(                                      #<2>
                                  str_extract(                                   #<2>
                                    cohort_description_num_paired_samples,       #<2>
                                    "\\d+$"                                      #<2>
                                    )                                            #<2>
                                ),                                               #<2>
           cohort_description_text = str_remove(                                 #<3>
                                        cohort_description_num_paired_samples,   #<3>
                                        "\\d+$"                                  #<3>
                                      ),                                         #<3>
           cohort_description = case_when(lead(is.na(dataset_name),              #<4>
                                              default = FALSE) ~                 #<4>
                                          paste0(cohort_description_text,        #<5>
                                                 lead(cohort_description_text)), #<5>
                                          .default = cohort_description_text)    #<6>
           ) %>%
    filter(!is.na(dataset_name)) %>%                                             #<7>
    select(!c(X4,                                                                #<8>
              cohort_description_num_paired_samples,                             #<8>
              cohort_description_text))                                          #<8>

```

1. Rename the columns so they are descriptive
2. Extracting the numbers (which represent the number of paired samples) at the end of the cohort descriptions to put them in their own column
3. Making a intermediate column, specifically a version of the text column that has those numbers of paired samples removed
4. Setting up a `case_when` that evaluates whether the next value/row in the dataset name column is an NA value. This utilizes the `lead` function to look for the next value. In the case of the last row, we're using the `default` argument to pad the data with a `FALSE`
5. If the `lead` function returns that the next value is NA (`TRUE`), it will paste the current cohort text with the next row's text
6. Otherwise (e.g., if the `lead` function returns that the next value is not NA (`FALSE`) or is the last value/row), it will return the current cohort text only.
7. This filters to remove the rows where the dataset name is NA because we've already appended the important information from those rows to the previous rows
8. This removes the `X4` column which was all `NAs`, the original third column (that had the cohort description plus number of paired samples), and the intermediate column which we made that has incomplete cohort descriptions

If we were to inspect the `table1_extract` Data Frame, we would observe that the data now appears to match our expectations. There are 14 rows and 7 columns, though those columns don't follow the original order of Table 1 in @Muller_Algavi_Borenstein_2022. Since the columns have descriptive names though, it will be fine that they are not in the same order. While the columns may not be in the same order as the source data, the rows should be in the same order as the source data.

```{r wrangled_data, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = 'The wrangled data appears to match our expectations.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f5db0d7ede_0_0#slide=id.g3f5db0d7ede_0_0")
```
If running code in a container in a non-interactive way (e.g., without an IDE) we could utilize output messages within our code to report on things like the dimensions of the Data Frame and to spot check specific values.

As an example, the following two messages report first on the dimensions and then spot check the dataset name for the fourth row of data.

```{r, eval = FALSE}
message("Dimensions: ", paste(dim(table1_extract), collapse = " x "))
message(table1_extract[4, "dataset_name"])
```

The dimensions properly report a decrease in the number of rows compared to the raw extracted data (14 instead of 25), matching our expectation for both the number of rows and the number of columns given the source data. The dataset name in the fourth row is `HE_INFANTS_MFGM_2019`, again matching our expectations when comparing to the source data.

To allow users to skip import and wrangling we will save the data as an RDA file as well as a CSV file as this is often useful to send our data to collaborators. We will save this in a “wrangled” subdirectory of our “data” directory of our working directory.

```{r, eval = FALSE}
save(table1_extract, file = here::here("data", "wrangled",       # <1>
                                       "wrangled_data.rda"))     # <1>
readr::write_csv(table1_extract,                                 # <2>
                 file = here::here("data","wrangled",            # <2>
                                   "wrangled_data.csv"))         # <2>
```

1. saving the data as an RDA file within the wrangled data subdirectory
2. saving the data as a CSV file within the wrangled data subdirectory

<details><summary>A more generalizable and robust wrangling approach</summary>

This wrangling approach using the `lead` function to assess if there is multiline data that needs combined assumes that there are only ever 1 - 2 rows for each cohort description. What changes to this process could be utilized that would be more generalizable or robust with data that doesn't meet this assumption? Specifically, how could this process be changed to combine an unknown number of rows of data to the previous row when only one column has data?

<details><summary>A possible approach</summary>

One possible approach would:

1. Use the `cumsum()` function to assign a group ID to the rows.
2. Use the `group_by()` function to group the data by that group ID.
3. With the `summarize()` function, `paste()` the cohort descriptions within each group together.
4. Either bring along the non-NA data for the other columns within the `summarize()` function, or join the datasets together using the group ID and drop unnecessary rows.

</details>

</details>


Now that we've wrangled the extracted table data into a Data Frame, we want to plan out the visualizations we'll make using this Data Frame!

Collab O. Rator's email provided the following guidance regarding the data visualizations:

>Specifically, I’d like to see scatter plots comparing the data from the final four columns for each dataset (No. samples w/ paired data, Longitudinal, No. HMDB-annotated compounds, and No. KEGG-annotated compounds).

Collab O. Rator's instructions have asked us to visualize 4 variables or dimensions of the data with scatter plots. The axes (x- and y-) will account for two of those variables, and we can adjust the points (e.g., color or shape) to represent a third variable -- likely the `longitudinal` value as this is categorical ("Yes" vs "No"). To visualize the fourth variable, we would need to add a third axis (z-) or some other means of encoding the data, but this adds to the complexity of the plot. Instead, we could make two separate panels.

Plot element | Data column
-------------|------------
x-axis       | `num_paired_samples`
y-axis       | 'hmdb_annotated_compounds' or `kegg_annotated_compounds`
point color or shape | `longitudinal`
labels               | `dataset_name`

But if we make two separate scatter plots, they will be two separate plots/files that won't be displayed next to each other unless we use design software like Illustrator to combine them later. Certain R packages allow you to combine plots and display them with each other. So let's "wrangle" the Dockerfile to include a package that will let us do that: `patchwork`.

## Container

As you can see from the visualization planning above, these two plots would be completely separate -- and we'd like them to be visible next to each other. A useful R package for this purpose is `patchwork`. However, our data visualization Docker image doesn't include that package.

Activity: Add `patchwork` to the container

* edit the Dockerfile (in your favorite text editor)
* build the Dockerfile to be an image (will need to do this on terminal)

***
