# **What are the data?**
***

The data comes from Table 1 in @Muller_Algavi_Borenstein_2022. Table 1 describes the curated gut microbiome-metabolome datasets that this study aggregated in order to curate a resource for the research community. This table is a 14 row 7 column table with columns showing

* the name of each dataset
* which reference in the study relates to that data
* a brief description of the cohort within that study
* the number of samples with paired fecal microbiome-metabolome data within that cohort
* whether the study contained longitudinal data
* the number of HMDB annotated compounds
* the number of KEGG annotated compounds

Each row in the table is providing data for that source dataset itself. Figure 1 in @Muller_Algavi_Borenstein_2022 provides more information on the overlap between datasets (e.g., in how many datasets were observed HMDB annotated metabolites present).

```{r data_table, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = 'Table 1 from Muller et al 2022. Table 1 describes the curated gut microbiome-metabolome datasets that this study collected and analyzed as part of the meta-analysis. This table is a 14 row 7 column table with columns showing the name of each dataset, which reference in the study relates to that data, a brief description of the cohort within that study, the number of samples with paired data within that cohort, whether the study utilized longitudinal data, as well as the number of HMDB annotated compounds and finally the number of KEGG annotated compounds in each study.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f5db0d7ede_0_25#slide=id.g3f5db0d7ede_0_25")
```

After using `tabulapdf` to extract the table from the PDF (see the `Data Import` section) and some basic wrangling with `tidyverse` (see the `Data Wrangling` section), we will extract a tidy dataframe with all of the columns displayed in the original table, but those boxed in purple and listed below are the most important for our visualization task (as outlined in Collab O. Rator's email).

```{r data_table_highlight, echo = FALSE, out.width = "100%", fig.align = 'center', fig.alt = 'The dataset name, number of paired samples, whether the sample is longitudinal, the number of HMDB annotated compounds, and the number of KEGG annotated compounds are the columns that we will use in building data visualizations.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f5db0d7ede_0_46#slide=id.g3f5db0d7ede_0_46")
```

Variable   | Details
---------- |-------------
**`dataset_name`**  | A character/string identifying the dataset (often containing author, studied illness, and year of the study).
**`num_paired_samples`**  | The number of samples within the dataset that have paired fecal microbiome-metabolome data.
**`longitudinal`** | A character/string reporting whether that dataset has longitudinal data ("Yes" or "No").
**`hmdb_annotated_compounds`** | The number of metabolites that were identified within that dataset and annotated with the Human Metabolome Database (HMDB) [@Wishart_etal_2009].
**`kegg_annotated_compounds`** | The number of metabolites that were identified within that dataset and annotated with Kyoto Encyclopedia of Genes and Genomes (KEGG) [@Kanehisa_Goto_2000; @Kanehisa_etal_2025].

***
