# **Data Import**
***

## Container

Activity: access the container the collaborator shared with you (because this is how you'll be able to get the data extracted from the PDF import).

* pull the container from Docker Hub

The other option is that we can build a Dockerfile but since we'll do that in a later activity, we'd rather show pulling from Docker Hub in this example.

Side quest: find a container on Docker Hub (and discuss reputable sources)

Note that although we can see the container (specifically an image at this step)-- this doesn't give us access to the data analysis data yet. For that we must run the container.

## Revisions task

Using the `tabulapdf` package, we will extract a raw, unwrangled version of Table 1.

Walk through how to run the container in order to run the analysis and extract the data.

Side quest: what if you wanted to spin up an IDE with your container?

Activity -- Using the container from the collaborator to get the Table 1 data.

Because we want to avoid mounting data and scripts for most of this case study (and leave that as part of the Data Analysis section or a side quest), we'll use the following instructions for someone to interact with the data analysis data within a running container:

For Docker Desktop:

* run the container
* navigate to the exec tab
* set-up a file structure (motivate this by it being important to have a file structure in a data analysis, pointing to ITCR reproducibility resources):
  * `mkdir opt/case_study_analysis plots/ data/imported data/wrangled`
* `touch` an `preprocessing.R` script in `opt/case_study_analysis`
* navigate to the files tab
* right click on the `preprocessing.R` script and click edit
* copy paste the following code into the script
* click save
* navigate to the exec tab
* use `Rscript preprocessing.R` -- adjusting based on script location and learner's file location.
* notice that it outputs "Dimensions: 25 x 7"
* notice that there is an rda file in the `data/imported` folder (which we can `ls data/imported` to see)

```{r, eval = FALSE}
#!/usr/bin/env Rscript

raw_table <- tabulapdf::extract_tables("https://www.nature.com/articles/s41522-022-00345-5.pdf",
                                       pages = 2,
                                       method = "stream",
                                       col_names = FALSE
                                      )[[1]]

message("Dimensions: ", paste(dim(raw_table), collapse = " x "))
```

```{r, eval = FALSE}
save(raw_table, file = here::here("data", "imported", "raw_table1.rda"))
```

***
