# **Summary**

***

## **Synopsis**

***

In this case study, you’ve learned about why containerization is used as part of reproducible science workflows. You’ve been introduced to Docker, and learned procedures for creating and modifying containers. You’ve practiced using the "Dockerfile -(build)-> image -(mount and run)-> container" process to create an active, isolated environment where you can perform an analysis. For your analysis, you pulled an image and created a container from it, running scripts and inspecting output within the container; and you modified the Dockerfile to contain a new package, built a new image, created another container and ran additional scripts, inspecting and saving the output from those. Finally, you created a script to drive the whole analysis and reviewed what needed to be shared with a collaborator or team member so they could validate the process.

Along the way, you’ve also learned about the advantages and challenges of containerization as well as secondary data resources and best practices for scientific project organization. We've seen that microbiome and metabolome data are combined to better understand the actual impact, not just the potential of the microorganisms within a microbiome. You've also learned that, in general, the greater the number of samples in a study, the higher the statistical power of a study, and more metabolites that are expected to be observed throughout the study.

You've learned that containers aren't always necessary and in some cases might not be worth the hassle, but in other cases they are beneficial and worth the challenge to learn and incorporate into your work. Your needs may go beyond what's covered in this resource which focused on using local containerization to run an analysis where the scripts were located in and results were saved to a mounted volume. Beyond the decision of whether containers are the right choice for you and your downstream use/work, working with containers requires decisions throughout the process.

```{r, echo = FALSE, fig.align = 'center', out.width = "100%", fig.alt = "Beyond the decision of whether containers are the right choice for you and your downstream use/work, working within the containerization process requires decisions: what you need available in your container; what can be shared and where; what files you'll need to mount (if any) when you run the container; whether you're running the container locally or remotely. This case study focuses on using containers to run an analysis locally (the bolded options), working with all parts of the containerization process."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1EFnuBBFlqjod7S1OlOoj_9-29nPO3ZhL6iJ-Pf8xzpg/edit?slide=id.g3f98af6272b_0_12#slide=id.g3f98af6272b_0_12")
```

## **Summary Plot**

***


```{r echo=FALSE, warning = FALSE, out.width=800, fig.alt = "Containers enhance reproducibility and scientific collaboration by providing a shareable and consistent computing environment. Multiple people in a team can contribute the software that they need for their part of the analysis into a visualization software container. This container can be stored on Docker Hub so they all can access it. Then the validator in the team can replicate everyone's work (perhaps data visualizations) using the same container without having to individually install the software from each team member."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Y2UAXo61mDtFkfToWYTIzcGzGfLzbpQ79I9ccD05A8c/edit?slide=id.p#slide=id.p")
```

***
